---
title: "cori.data.bfs"
description: "Access and analyze U.S. Census Bureau Business Formation Statistics (BFS) at county, state, and national level from 2005 to present"
author: "Drew Rosebush"
date: "2026-05-12"
categories: ["R Package", "Entrepreneurship"]
tags: ["BFS", "Business Formation", "Business Applications", "County-Level", "S3", "Tidy Data"]
package:
  name: "cori.data.bfs"
  githubUrl: "https://github.com/ruralinnovation/cori.data.bfs"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.bfs")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["census-bfs"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.bfs` package provides access to U.S. Census Bureau
Business Formation Statistics (BFS) at the county, state, and national
level from 2005 to present. BFS data tracks the number of business
applications filed with the IRS — a leading indicator of entrepreneurial
activity that precedes actual business formation by 6–18 months.

Data are returned in long format: one row per `geoid / year / variable`.

::: callout-note
## Why Use This Package

-   **Leading indicator** — Business applications precede actual firm
    births, making BFS the earliest available signal of entrepreneurship
    activity
-   **S3-backed access** — Pre-processed data available via DuckDB
    without downloading full Census source files
-   **Long panel** — Annual data from 2005 to present, refreshed each
    year after Census release
-   **Consistent geoids** — Standard 5-digit FIPS for counties, 2-digit
    for states, `"00"` for national
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.bfs")
```

## Quick Start

```{r}
#| eval: false

library(cori.data.bfs)

# All data, latest vintage
df <- read_bfs_from_s3()

# Business applications for specific counties and years
df <- read_bfs_from_s3(
  variables = "business_applications",
  geoids    = c("54011", "54025"),
  years     = 2010:2024
)

# Variable documentation
get_bfs_codebook()
```

## Key Features

### S3-Backed Data Access

Pre-processed data is stored in S3 and queried via DuckDB, returning
only the rows you need without downloading the full dataset:

```{r}
#| eval: false

# Filter by geography, variables, and year range
read_bfs_from_s3(
  variables = "business_applications",
  geoids    = c("54011"),
  years     = 2015:2024
)
```

### Data Refresh (Admin)

After Census releases new estimates (typically mid-year), refresh the
S3 store:

```{r}
#| eval: false

write_bfs_processed_to_s3(years = 2005:2024)
```

## Core Functions

| Function | Category | Description |
|---|---|---|
| `read_bfs_from_s3()` | Data Access | Read processed BFS data from S3 via DuckDB |
| `pull_bfs()` | Data Access | Pull raw data from Census source files |
| `write_bfs_processed_to_s3()` | Admin | Process and write BFS data to S3 |
| `get_bfs_codebook()` | Utility | Return variable definitions and metadata |
| `latest_bfs_vintage()` | Utility | Return the most recent available vintage year |

## Data Coverage

| Dimension | Coverage |
|---|---|
| **Variables** | `business_applications` (BA series) |
| **Geography** | National (`"00"`), State (2-digit FIPS), County (5-digit FIPS) |
| **Time Period** | 2005–present |
| **Update Frequency** | Annual (~6 months after year-end) |
| **Source File** | `bfs_county_apps_annual.xlsx` (Census Bureau) |

## Use Cases

-   **Leading indicator analysis** — Track business application trends
    6–18 months before firm births are recorded in BDS
-   **Rural entrepreneurship** — Compare rural vs. metro application
    rates at county level
-   **Economic development** — Identify counties with accelerating or
    decelerating startup formation
-   **Nowcasting** — Use latest BFS release to estimate upcoming BDS
    figures before annual publication

## Related Resources

### Datasets

-   [Census Business Formation Statistics](/datasets/census-bfs/) —
    Source dataset documentation
-   [Census Business Dynamics Statistics](/datasets/census-bds/) —
    Lagged companion measuring actual firm entry/exit

### Related Packages

-   [cori.data.bds](/packages/cori-data-bds/) — S3-backed access layer
    for Business Dynamics Statistics (actual firm births/deaths)

### External Links

-   [Census BFS Homepage](https://www.census.gov/econ/bfs/)
-   [BFS Methodology](https://www.census.gov/econ/bfs/methodology.html)
-   [Source File](https://www.census.gov/econ/bfs/xlsx/bfs_county_apps_annual.xlsx)

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data.bfs)
-   [Full README](https://github.com/ruralinnovation/cori.data.bfs/blob/main/README.md)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.bfs/issues).
