---
title: cori.data.bfs
description: Access and analyze U.S. Census Bureau Business Formation Statistics
  (BFS) at the county, state, and national level from 2005 to present. Measures
  business applications — EIN filings — as a leading indicator of entrepreneurial
  activity in rural and nonrural communities.
author: Drew Rosebush
date: '2026-08-31'
categories:
  - R Package
  - Entrepreneurship
tags:
  - BFS
  - Business Formation
  - EIN
  - County-Level
  - S3
  - Tidy Data
  - Rural
package:
  name: cori.data.bfs
  githubUrl: https://github.com/ruralinnovation/cori.data.bfs
  installCommand: remotes::install_github("ruralinnovation/cori.data.bfs")
  status: development
  version: 0.1.0
  maintainer: Drew Rosebush
  featured: false
usesDatasets:
  - census-bfs
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.bfs` package provides access to U.S. Census Bureau
Business Formation Statistics (BFS) at the county, state, and national
level from 2005 to the present. BFS measures business applications —
Employer Identification Number (EIN) filings with the IRS — as a
leading indicator of entrepreneurial activity. Because applications
precede establishment by months or years, the data captures
entrepreneurial momentum before it shows up in employment or business
counts.

Because applications precede establishments by months or years, BFS is
the earliest observable signal of whether entrepreneurial momentum is
building in a community — particularly useful for tracking rural
economies where establishment counts alone may show little movement. All
data are returned in the standard `cori.data` long format: one row per
`geoid / year / variable`.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/cori.data.bfs")
```

```{r}
#| eval: false
library(cori.data.bfs)

# All variables and their definitions
get_bfs_codebook()

# Total business applications — all counties, all years
get_business_applications(geography = "county")

# New Hampshire counties, 2010 onward
get_business_applications(
  geoids = c("33009", "33011"),
  years  = 2010:2024
)
```

## What This Data Tells Us About Rural

Business applications are best interpreted as a leading indicator of
future establishment counts rather than a measure of current economic
strength. Not all applications convert to employer businesses, and the
conversion rate varies by community conditions, industry, and access to
capital and support. Reading application trends alongside employment and
wage data gives the fullest picture of whether entrepreneurial momentum
is translating into durable economic change.

## Introduction

*Adapted from the [Introduction
vignette](https://github.com/ruralinnovation/cori.data.bfs/blob/main/vignettes/introduction.Rmd)
on GitHub.*

`cori.data.bfs` provides annual business application counts from the
U.S. Census Bureau's Business Formation Statistics. Data cover
2005–present and measure total EIN
applications filed with the IRS, excluding tax liens, estates, trusts,
agriculture (NAICS 11), and public administration (NAICS 92).

**Source:** U.S. Census Bureau, Business Formation Statistics\
**Coverage:** 2005–present, updated annually\
**Geography:** County (5-digit FIPS), state (2-digit FIPS), national (`"00"`)

### Variables

```{r}
#| eval: false
get_bfs_codebook() |>
  dplyr::select(variable, label, unit, notes) |>
  knitr::kable()
```

### Reading data

All data are returned in long format: one row per `geoid / year /
variable`.

```{r}
#| eval: false
df <- get_business_applications(geography = "county")
dplyr::glimpse(df)
```

Filter to specific geographies, years, or variables:

```{r}
#| eval: false
nh_bfs <- get_business_applications(
  geoids = grep("^33", unique(df$geoid), value = TRUE),
  years  = 2010:2024
)
dplyr::glimpse(nh_bfs)
```

### Rural vs. Nonrural

Business formation diverges between rural and nonrural counties. The
chart below shows total business applications by rural status using the
CBSA 2023 rural definition.

```{r}
#| eval: false
library(cori.charts)
library(ggplot2)
library(ruraldefinitions)
library(dplyr)

load_fonts()

rural_xwalk <- ruraldefinitions::cbsa_2023 |> select(geoid, is_rural)

rural_avg <- df |>
  left_join(rural_xwalk, by = "geoid") |>
  filter(!is.na(is_rural)) |>
  group_by(year, is_rural) |>
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop")

ggplot(rural_avg, aes(x = year, y = value, color = is_rural)) +
  annotate("rect", xmin = 2007.9, xmax = 2009.5,
           ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.4) +
  annotate("rect", xmin = 2020, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.4) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_color_manual(
    values = c("Rural" = "#2F6E9B", "Nonrural" = "#7EBDC2"),
    labels = c("Rural" = "Rural counties", "Nonrural" = "Nonrural counties")
  ) +
  scale_x_continuous(
    breaks = seq(2005, 2024, by = 4),
    expand = expansion(mult = c(0, 0.2))
  ) +
  scale_y_continuous(labels = scales::label_number(scale = 1e-3, suffix = "K")) +
  theme_cori() +
  theme(legend.position = "bottom") +
  labs(
    title    = "Rural business formation surged during the pandemic and has held",
    subtitle = "Total business applications (BA series), 2005\u20132024",
    x        = NULL,
    y        = NULL,
    color    = NULL,
    caption  = paste0(
      "Source: CORI analysis of U.S. Census Bureau Business Formation Statistics.\n",
      "Shaded bands mark Great Recession (2008\u201309) and COVID-19 (2020). ",
      "Rural classification: CORI CBSA 2023 definition."
    )
  )
```

### County Spotlight: Grafton County, NH

```{r}
#| eval: false
library(cori.data.pep)

grafton_bfs <- get_business_applications(
  geoids = c("33009", "33", "00"),
  years  = 2005:2024
)

grafton_pop <- get_population(
  geoids    = c("33009", "33", "00"),
  variables = "population",
  years     = 2005:2024
) |>
  select(geoid, year, population = value)

grafton <- grafton_bfs |>
  left_join(grafton_pop, by = c("geoid", "year")) |>
  mutate(
    apps_per_1k = value / population * 1000,
    group = case_when(
      geoid == "33009" ~ "Grafton County, NH",
      geoid == "33"    ~ "New Hampshire",
      geoid == "00"    ~ "United States"
    ),
    group = factor(group, levels = c("Grafton County, NH", "New Hampshire", "United States"))
  )

ggplot(grafton, aes(x = year, y = apps_per_1k,
                    color = group, linetype = group, linewidth = group)) +
  annotate("rect", xmin = 2007.9, xmax = 2009.5,
           ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.4) +
  annotate("rect", xmin = 2020, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.4) +
  geom_line(na.rm = TRUE) +
  geom_point(size = 2.5, na.rm = TRUE) +
  scale_color_manual(values = c(
    "Grafton County, NH" = "#2F6E9B",
    "New Hampshire"      = "#7EBDC2",
    "United States"      = "#9DA7B0"
  )) +
  scale_linetype_manual(values = c(
    "Grafton County, NH" = "solid",
    "New Hampshire"      = "dashed",
    "United States"      = "dotted"
  )) +
  scale_linewidth_manual(values = c(
    "Grafton County, NH" = 1.2,
    "New Hampshire"      = 0.8,
    "United States"      = 0.8
  )) +
  scale_x_continuous(breaks = seq(2005, 2024, by = 4)) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
  theme_cori() +
  theme(legend.position = "bottom") +
  labs(
    title    = "Business applications per 1,000 residents \u2014 Grafton County, NH",
    subtitle = "Compared to New Hampshire and the United States, 2005\u20132024",
    x        = NULL, y = NULL, color = NULL, linetype = NULL, linewidth = NULL,
    caption  = paste0(
      "Source: CORI analysis of U.S. Census Bureau Business Formation Statistics.\n",
      "Population denominator from Census Population Estimates Program via cori.data.pep.\n",
      "Shaded bands mark Great Recession (2008\u201309) and COVID-19 (2020)."
    )
  )
```

## Related

-   **Dataset:** [Census Business Formation Statistics](/datasets/census-bfs/)
-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `cori.data.bfs`
-   **Package:** [cori.data.bds](/packages/cori-data-bds/) — establishment
    counts and job creation; pairs with BFS for the full formation-to-employment story
-   **Package:** [cori.data.pep](/packages/cori-data-pep/) — population
    denominators used in the county spotlight above
-   **Project:** [Rural Economic
    Outlook](/projects/rural-economic-outlook/) — BFS is a core input to
    CORI's multi-dataset economic analysis

::: callout-note
**Data attribution:** Business Formation Statistics are produced by the
U.S. Census Bureau. When citing analysis using this package, reference:
U.S. Census Bureau, Business Formation Statistics,
[www.census.gov/econ/bfs](https://www.census.gov/econ/bfs).
:::
