---
title: "cori.data.bfs"
description: "Access and analyze U.S. Census Bureau Business Formation Statistics (BFS) at county, state, and national level from 2005 to present"
author: "Drew Rosebush"
date: "2026-08-26"
categories: ["R Package", "Entrepreneurship"]
tags: ["BFS", "Business Formation", "Business Applications", "County-Level", "S3", "Tidy Data"]
package:
  name: "cori.data.bfs"
  githubUrl: "https://github.com/ruralinnovation/cori.data.bfs"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.bfs")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["census-bfs"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.bfs` package provides access to U.S. Census Bureau
Business Formation Statistics (BFS) at the county, state, and national
level from 2005 to present. BFS measures total EIN applications filed
with the IRS — excluding tax liens, estates, trusts, agriculture
(NAICS 11), and public administration (NAICS 92) — a leading indicator
of entrepreneurial activity that precedes actual business formation by
6–18 months.

Data are returned in long format: one row per `geoid / year / variable`.

::: callout-note
## What's in the package

-   **Leading indicator** — Business applications precede actual firm
    births, making BFS the earliest available signal of entrepreneurship
    activity
-   **S3-backed access** — Pre-processed data available via DuckDB
    without downloading full Census source files
-   **Long panel** — Annual data from 2005 to present, refreshed each
    year after Census release
-   **Consistent geoids** — 5-digit FIPS for counties, 2-digit for
    states, `"00"` for national
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.bfs")
```

## Quick start

```{r}
#| eval: false

library(cori.data.bfs)

# All data, latest vintage
df <- read_bfs_from_s3()

# Business applications for specific counties and years
df <- read_bfs_from_s3(
  variables = "business_applications",
  geoids    = c("54011", "54025"),
  years     = 2010:2024
)

# Variable documentation
get_bfs_codebook()
```

## Core functions

### Reading data

`read_bfs_from_s3()` reads pre-processed data straight from S3 via
DuckDB — filter by geography, variables, and year range without
downloading the full dataset:

```{r}
#| eval: false

read_bfs_from_s3(
  variables = "business_applications",
  geoids    = c("54011"),
  years     = 2015:2024
)
```

`get_business_applications()` is the tidy-format entry point used
throughout analysis workflows:

```{r}
#| eval: false

df <- get_business_applications(geography = "county")

# Filter to specific geographies and years
nh_bfs <- get_business_applications(
  geoids = grep("^33", unique(df$geoid), value = TRUE),
  years  = 2010:2024
)
```

### Data refresh (admin)

After Census releases new estimates (typically ~6 months after
year-end), refresh the S3 store:

```{r}
#| eval: false

write_bfs_processed_to_s3(years = 2005:2024)
```

## Key variables

| Variable | Description |
|---|---|
| `business_applications` | Total business applications (BA series) |

| Geography | geoid format | Example |
|---|---|---|
| County | 5-digit FIPS | `"33009"` |
| State | 2-digit FIPS | `"33"` |
| National | `"00"` | `"00"` |

Use `get_bfs_codebook()` for complete variable definitions, units, and notes.

## Common use cases

### Rural vs. nonrural business formation

Business formation diverges between rural and nonrural counties.
Joining BFS to population data (via `cori.data.pep`) and CORI's rural
definitions produces a per-capita comparison:

```{r}
#| eval: false

library(cori.data.pep)
library(ruraldefinitions)
library(dplyr)

rural_definitions <- cbsa_2023 |> select(geoid, is_rural)

county_pop <- get_population(
  geoids    = unique(df$geoid),
  variables = "population",
  years     = 2005:2024
) |>
  select(geoid, year, population = value)

rural_avg <- df |>
  left_join(rural_definitions, by = "geoid") |>
  left_join(county_pop, by = c("geoid", "year")) |>
  filter(!is.na(is_rural), !is.na(population), population > 0) |>
  group_by(year, is_rural) |>
  summarise(
    applications = sum(value, na.rm = TRUE),
    population   = sum(population, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(apps_per_1k = applications / population * 1000)
```

CORI's own analysis using this pattern found rural business formation
surged during the pandemic and has remained elevated relative to its
pre-2020 trend.

### County spotlight

The same `geoids` filter works across geography levels in a single
call, useful for comparing a county against its state and the nation:

```{r}
#| eval: false

grafton_bfs <- get_business_applications(
  geoids = c("33009", "33", "00"),  # county, state, national
  years  = 2005:2024
)
```

## Data coverage

| Dimension | Coverage |
|---|---|
| **Variables** | `business_applications` (BA series) |
| **Geography** | National (`"00"`), state (2-digit FIPS), county (5-digit FIPS) |
| **Time Period** | 2005–present |
| **Update Frequency** | Annual (~6 months after year-end) |
| **Source File** | `bfs_county_apps_annual.xlsx` (Census Bureau) |

## Related resources

### Datasets

-   [Census Business Formation Statistics](/datasets/census-bfs/) —
    Source dataset documentation
-   [Census Business Dynamics Statistics](/datasets/census-bds/) —
    Lagged companion measuring actual firm entry/exit

### Related packages

-   [cori.data.bds](/packages/cori-data-bds/) — Business Dynamics
    Statistics (actual firm births/deaths)
-   [cori.data.pep](/packages/cori-data-pep/) — Population estimates,
    used as the denominator for per-capita comparisons

### External links

-   [Census BFS Homepage](https://www.census.gov/econ/bfs/)
-   [BFS Methodology](https://www.census.gov/econ/bfs/methodology.html)
-   [Source File](https://www.census.gov/econ/bfs/xlsx/bfs_county_apps_annual.xlsx)

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data.bfs)
-   [Full README](https://github.com/ruralinnovation/cori.data.bfs/blob/main/README.md)
-   [Introduction Vignette](https://github.com/ruralinnovation/cori.data.bfs/blob/main/vignettes/introduction.Rmd)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.bfs/issues).
