---
title: cori.data.qcew
description: Access and analyze Bureau of Labor Statistics Quarterly Census of Employment
  and Wages (QCEW) data at the county, state, and national level from 1990 to present.
  Covers total employment, average annual pay, sector composition, and employment
  concentration using CORI's tradable goods / tradable services / local services
  classification.
author: Drew Rosebush
date: '2026-08-31'
categories:
  - R Package
  - Labor Market
tags:
  - QCEW
  - BLS
  - Employment
  - Wages
  - County-Level
  - S3
  - Tidy Data
  - Rural
package:
  name: cori.data.qcew
  githubUrl: https://github.com/ruralinnovation/cori.data.qcew
  installCommand: remotes::install_github("ruralinnovation/cori.data.qcew")
  status: development
  version: 0.1.0
  maintainer: Drew Rosebush
  featured: false
usesDatasets:
  - qcew-employment-wages
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.qcew` package provides access to Bureau of Labor
Statistics Quarterly Census of Employment and Wages (QCEW) data at the
county, state, and national level from 1990 to the present. Derived from
unemployment insurance records rather than a survey, QCEW covers roughly
95% of all U.S. employment — the standard source for local labor market
analysis, and more reliable than survey-based estimates in the small
rural counties where sample sizes are thin.

Beyond access, the package adds two analytical layers over the raw
federal source. The first is CORI's supersector classification — a
three-way grouping of BLS's 11 NAICS supersectors into Tradable Goods,
Tradable Services, and Local Services, loosely following the methodology
in Eckert (2018). This classification distinguishes industries that bring
external income into a community from those that primarily serve local
demand, which is central to understanding a rural economy's long-run wage
trajectory. The second is an employment concentration measure (HHI) that
reveals how economically diversified a county is across sectors — a
structural indicator that employment headcounts alone cannot provide.

All functions return tidy long-format data consistent with the
`cori.data` ecosystem: `geoid | year | variable | value`, with an
`agg_var` employment-weight column on rate and share variables for
correct aggregation across geographies of different sizes.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/cori.data.qcew")
```

```{r}
#| eval: false
library(cori.data.qcew)

# See all available variables and which functions return them
get_qcew_codebook()

# Total employment — all counties, all years
get_employment(geography = "county")

# Average annual pay — specific counties
get_wage_salary(geoids = c("33009", "33011"), years = 2010:2023)

# Employment by CORI supersector
get_sector_employment(geography = "county", years = 2018:2023,
                      sector_type = "CORI", value_type = "share")

# Employment concentration (HHI)
get_employment_concentration(geography = "county", years = 2023)
```

Check the current data vintage:

```{r}
#| eval: false
latest_qcew_vintage()
```

## What This Data Tells Us About Rural

Employment is the most fundamental indicator of a local economy's health
— and one of the most easily misread in a rural context. Aggregate
employment counts can grow while wages stagnate, or decline while
productivity rises. In CORI's community research, many rural counties
show GDP per capita growth alongside employment that has not recovered to
prior peaks — a pattern that reflects a smaller, more productive
workforce rather than a stagnant one. The headline number and the
underlying story can point in different directions.

What the QCEW data adds beyond simple employment counts is sector
composition. CORI's three-way classification — tradable goods, tradable
services, and local services — is particularly useful for rural analysis.
Local services (retail, healthcare, food service) meet local demand and
scale with population; they are essential to community function but do
not drive wage growth or export earnings. Tradable goods and services —
manufacturing, agriculture, finance, tech, professional services — bring
external income into a community and shape its long-run wage trajectory.
Rural counties with strong tradable sectors reflect economies whose
historical asset base remains economically active. Those shifting toward
tradable services are often communities building new competitive
advantages alongside inherited ones.

Industry concentration is a structural feature of many rural economies,
reflecting the historic dominance of particular assets — a major
employer, an agricultural commodity, a natural resource. Concentration
can be a strength (scale, specialization, deep supply chains) and a
source of vulnerability (exposure if that industry contracts). The QCEW
data makes this composition visible at the county level and trackable
over time, supporting both the identification of strengths and the
analysis of diversification opportunities.

Average annual pay reflects both the composition of employment and wages
within sectors. Understanding the wage structure of a rural economy —
not just its size — is essential for reading its trajectory and its
capacity to retain talent.

## Introduction

*Adapted from the [Introduction
vignette](https://github.com/ruralinnovation/cori.data.qcew/blob/main/vignettes/introduction.Rmd)
on GitHub.*

### What is this data?

The Bureau of Labor Statistics Quarterly Census of Employment and Wages
(QCEW) covers virtually all U.S. jobs — roughly 95% of all employment
— through the unemployment insurance system. Unlike surveys, QCEW is a
near-census of covered employment and wages, making it the standard
source for local labor market analysis.

Data is available at the **county and state level** from **1990 to the
present**.

### Sector breakdowns

Two sector classification systems are available via `sector_type`:

-   **`"BLS"`** (default) — 11 NAICS supersectors from the source data
-   **`"CORI"`** — 3 custom supersectors: Tradable Goods, Tradable
    Services, and Local Services

```{r}
#| eval: false
# BLS sector employment counts
bls_emp <- get_sector_employment(
  geography   = "county",
  years       = 2018:2023,
  sector_type = "BLS",
  value_type  = "count"
)

# CORI super-sector employment shares
cori_share <- get_sector_employment(
  geography   = "county",
  years       = 2018:2023,
  sector_type = "CORI",
  value_type  = "share"
)
```

```{r}
#| eval: false
# Average pay by CORI super-sector
cori_pay <- get_sector_wages(
  geography   = "county",
  years       = 2018:2023,
  sector_type = "CORI"
)
```

**Note on CORI counts:** CORI supersector employment counts are not
stored directly. To derive approximate counts from shares, multiply
`value` by `agg_var`:

```{r}
#| eval: false
library(dplyr)

cori_share |>
  mutate(approx_employment = value * agg_var)
```

### Filtering by geography

```{r}
#| eval: false
# Specific counties
emp <- get_employment(geoids = c("33009", "33011"), years = 2010:2023)

# Specific states
pay <- get_wage_salary(geoids = c("33", "50", "23"), years = 2015:2023)
```

### Rural vs. non-rural comparison

```{r}
#| eval: false
library(dplyr)

county_emp <- get_employment(geography = "county", years = 2010:2023)

rural_definition <- ruraldefinitions::cbsa_2023 |> select(geoid, is_rural)

county_comparison <- county_emp |>
  left_join(rural_definition, by = "geoid") |>
  filter(!is.na(is_rural), year == 2023) |>
  group_by(year, is_rural) |>
  summarize(total_employment = sum(value, na.rm = TRUE), .groups = "drop")

print(county_comparison)
```

### Understanding agg_var

Functions that return rates or shares include an `agg_var` column — the
employment weight used for proper aggregation across geographies. To
compute an employment-weighted average pay across a set of counties:

```{r}
#| eval: false
pay <- get_wage_salary(geography = "county", years = 2023)

pay |>
  summarize(
    weighted_avg_pay = sum(value * agg_var, na.rm = TRUE) /
                       sum(agg_var, na.rm = TRUE)
  )
```

## Rural vs. Nonrural Employment Trends

*Adapted from the [Rural vs. Nonrural Employment Trends
vignette](https://github.com/ruralinnovation/cori.data.qcew/blob/main/vignettes/rural-employment-trends.Rmd)
on GitHub.*

This vignette compares employment trends between rural and nonrural
counties since the Great Recession using `get_employment()`, the
`ruraldefinitions` package, and `cori.charts` for visualization.

```{r}
#| eval: false
library(cori.data.qcew)
library(cori.charts)
library(ruraldefinitions)
library(ggrepel)
library(dplyr)
library(ggplot2)

load_fonts()
```

### Rural vs. nonrural employment since the Great Recession

Pull county-level annual employment for all years since 2007:

```{r}
#| eval: false
emp_raw <- get_employment(
  geography = "county",
  years     = 2007:2024
)
```

Keep only 5-digit county geoids and attach the rural/nonrural flag from
`ruraldefinitions::cbsa_2023`:

```{r}
#| eval: false
emp_co <- emp_raw |>
  filter(nchar(geoid) == 5) |>
  left_join(
    cbsa_2023 |> select(geoid, is_rural),
    by = "geoid"
  )
```

Index employment to 2007 levels within each group to show relative
change:

```{r}
#| eval: false
chrt_dta <- emp_co |>
  filter(!is.na(is_rural)) |>
  group_by(is_rural, year) |>
  reframe(tot_emp = sum(value)) |>
  group_by(is_rural) |>
  mutate(
    index_value = tot_emp / tot_emp[year == 2007],
    year        = as.numeric(year)
  )
```

Quick stats for the most recent year-over-year change:

```{r}
#| eval: false
latest_yr <- max(chrt_dta$year)

chrt_dta |>
  filter(year >= (latest_yr - 1)) |>
  arrange(is_rural, year) |>
  group_by(is_rural) |>
  mutate(pct_change = (tot_emp - lag(tot_emp)) / lag(tot_emp)) |>
  filter(year == latest_yr) |>
  select(is_rural, pct_change)
```

```{r}
#| eval: false
label_df <- chrt_dta |> filter(year == max(year))

# NOTE: chart title uses "fragile" — revisit framing before publishing
fig_line <- chrt_dta |>
  ggplot(aes(x = year, y = index_value, color = is_rural, group = is_rural)) +
  annotate("rect", xmin = 2007, xmax = 2009, ymin = .9, ymax = 1.2,
           fill = "grey90", alpha = 0.5) +
  geom_line(linewidth = 1.5) +
  geom_text_repel(
    data          = label_df,
    aes(label     = is_rural),
    nudge_x       = 1.5,
    direction     = "y",
    hjust         = 0,
    segment.color = NA,
    size          = 5
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(.9, 1.2),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    expand   = expansion(mult = c(0, 0)),
    n.breaks = 5,
    labels   = scales::label_number(accuracy = 1)
  ) +
  scale_color_manual(
    values = c(
      "Rural"    = cori_palettes$ctg2ruralnonrural[[1]],
      "Nonrural" = cori_palettes$ctg2ruralnonrural[[2]]
    )
  ) +
  theme_cori() +
  theme(
    legend.position = "none",
    axis.ticks.x    = element_line(color = "#d0d2ce", linewidth = .25),
    plot.title      = element_text(color = cori_colors[["Dark Green"]])
  ) +
  labs(
    title    = "Rural employment set back twice since the Great Recession",
    subtitle = "Relative change in employment since 2007",
    x        = NULL,
    y        = NULL,
    caption  = paste0(
      "Source: CORI analysis of BLS Quarterly Census of Employment and Wages.\n",
      "Note: Indexed to 2007 levels. Rural = nonmetro counties per CBSA 2023 definition."
    )
  )

fig_line
```

### Regional breakdown

To see how rural employment trends differ by Census region:

```{r}
#| eval: false
region_xwalk <- tigris::states() |>
  sf::st_drop_geometry() |>
  select(GEOID, REGION) |>
  mutate(
    region = case_when(
      REGION == "1" ~ "Northeast",
      REGION == "2" ~ "Midwest",
      REGION == "3" ~ "South",
      REGION == "4" ~ "West"
    )
  )

regional_rural <- emp_co |>
  filter(!is.na(is_rural), is_rural == "Rural") |>
  mutate(st_fips = substr(geoid, 1, 2)) |>
  left_join(region_xwalk, by = c("st_fips" = "GEOID")) |>
  filter(!is.na(region)) |>
  group_by(region, year) |>
  reframe(tot_emp = sum(value)) |>
  group_by(region) |>
  mutate(
    index_value = tot_emp / tot_emp[year == 2007],
    year        = as.numeric(year)
  )
```

```{r}
#| eval: false
region_colors <- c(
  "West"      = cori_palettes$ctg4mid[[1]],
  "South"     = cori_palettes$ctg4mid[[2]],
  "Midwest"   = cori_palettes$ctg4mid[[3]],
  "Northeast" = cori_palettes$ctg4mid[[4]]
)

label_df <- regional_rural |> filter(year == max(year))

regional_rural |>
  ggplot(aes(x = year, y = index_value, color = region)) +
  annotate("rect", xmin = 2007, xmax = 2009, ymin = -Inf, ymax = Inf,
           fill = "grey90", alpha = 0.5) +
  geom_hline(yintercept = 1, color = "grey70", linetype = "dashed") +
  geom_line(linewidth = 1.5) +
  geom_text_repel(
    data          = label_df,
    aes(label     = region),
    nudge_x       = 0.5,
    direction     = "y",
    hjust         = 0,
    segment.color = NA,
    size          = 4.5,
    show.legend   = FALSE
  ) +
  scale_color_manual(values = region_colors) +
  scale_x_continuous(
    limits = c(2007, max(regional_rural$year) + 3),
    expand = expansion(mult = c(0.02, 0)),
    labels = scales::label_number(accuracy = 1, big.mark = "")
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    expand = expansion(mult = c(0, .1))
  ) +
  theme_cori() +
  theme(
    legend.position = "none",
    axis.ticks.x    = element_line(color = "#d0d2ce", linewidth = .25),
    plot.title      = element_text(color = cori_colors[["Dark Green"]])
  ) +
  labs(
    title    = "Rural employment by Census region: West leads recovery",
    subtitle = "Relative change in rural employment since the Great Recession",
    x        = NULL,
    y        = NULL,
    caption  = paste0(
      "Source: CORI analysis of BLS Quarterly Census of Employment and Wages.\n",
      "Note: Indexed to 2007 levels. Rural = nonmetro counties per CBSA 2023 definition."
    )
  )
```

### Industry concentration: HHI

The Herfindahl-Hirschman Index (`employment_hhi`) measures how
concentrated a county's employment is across industries. Higher values
indicate fewer industries dominating local employment — a sign of lower
diversification. Comparing rural and nonrural counties in the Midwest
illustrates how this plays out within a single region.

```{r}
#| eval: false
library(ggbeeswarm)

midwest_fips <- c("17", "18", "19", "20", "26", "27", "29",
                  "31", "38", "39", "46", "55")

emp_raw <- get_employment(geography = "county", years = 2024) |>
  filter(nchar(geoid) == 5, substr(geoid, 1, 2) %in% midwest_fips) |>
  select(geoid, employment = value)

hhi_raw <- get_employment_concentration(geography = "county", years = 2024) |>
  select(geoid, employment_hhi = value)

midwest_co <- emp_raw |>
  left_join(hhi_raw, by = "geoid") |>
  left_join(cbsa_2023 |> select(geoid, is_rural), by = "geoid") |>
  filter(!is.na(is_rural), !is.na(employment_hhi))
```

```{r}
#| eval: false
fig_swarm <- midwest_co |>
  ggplot(aes(x = employment_hhi, y = is_rural,
             size = employment, color = is_rural)) +
  geom_beeswarm(cex = 0.6, groupOnX = FALSE, alpha = 0.4) +
  scale_x_continuous(
    labels = scales::label_comma(),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_manual(
    values = c(
      "Rural"    = cori_palettes$ctg2ruralnonrural[[1]],
      "Nonrural" = cori_palettes$ctg2ruralnonrural[[2]]
    ),
    guide = "none"
  ) +
  scale_size(
    range  = c(0.5, 6),
    labels = scales::label_comma(scale_cut = scales::cut_short_scale()),
    guide  = guide_legend(
      title          = "Employment",
      title.position = "left",
      title.theme    = element_text(margin = margin(r = 8), size = 11),
      label.theme    = element_text(margin = margin(r = 5), size = 11)
    )
  ) +
  theme_cori() +
  labs(
    title    = "Rural Midwest counties have more concentrated economies",
    subtitle = "Employment HHI by rural status, Midwest counties (2024)",
    x        = "Employment HHI (higher = more concentrated)",
    y        = NULL,
    caption  = paste0(
      "Source: CORI analysis of BLS Quarterly Census of Employment and Wages.\n",
      "Note: Each dot is one county, sized by total employment. ",
      "Rural = nonmetro per CBSA 2023 definition."
    )
  )

fig_swarm
```

## Related

-   **Dataset:** [QCEW Employment &
    Wages](/datasets/qcew-employment-wages/) — source dataset
    documentation
-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `cori.data.qcew`
-   **Package:** [ruraldefinitions](/packages/ruraldefinitions/) —
    rural/nonrural county classifications used throughout
-   **Project:** [Rural Economic
    Outlook](/projects/rural-economic-outlook/) — QCEW is a core input
    to CORI's multi-dataset economic analysis

::: callout-note
**Data attribution:** QCEW data is produced by the U.S. Bureau of Labor
Statistics. When citing analysis using this package, reference: U.S.
Bureau of Labor Statistics, Quarterly Census of Employment and Wages,
[www.bls.gov/cew](https://www.bls.gov/cew/).
:::
