---
title: "cori.data.qcew"
description: "Access and analyze Bureau of Labor Statistics Quarterly Census of Employment and Wages (QCEW) data at the county and state level from 1990 to present"
author: "Drew Rosebush"
date: "2026-08-26"
categories: ["R Package", "Labor Market"]
tags: ["QCEW", "Employment", "Wages", "NAICS", "County-Level", "S3", "Tidy Data"]
package:
  name: "cori.data.qcew"
  githubUrl: "https://github.com/ruralinnovation/cori.data.qcew"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.qcew")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["qcew-employment-wages"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.qcew` package provides access to Bureau of Labor
Statistics Quarterly Census of Employment and Wages (QCEW) data at the
county and state level, covering 1990 to the present. QCEW covers
roughly 95% of all U.S. jobs through the unemployment insurance
system — unlike a survey, it's a near-census of covered employment and
wages, making it the standard source for local labor market analysis.

::: callout-note
## What's in the package

-   **Five analysis functions** — total employment, average pay,
    sector employment, sector wages, and employment concentration (HHI)
-   **Two sector taxonomies** — BLS's 11 source-level supersectors, or
    CORI's own 3-way Tradable Goods / Tradable Services / Local
    Services classification
-   **Weighted aggregation built in** — rate and share variables carry
    an `agg_var` employment-weight column
-   **Long panel** — annual data from 1990 to present, always
    pointing at the latest vintage
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.qcew")
```

## Quick start

```{r}
#| eval: false

library(cori.data.qcew)

# See all available variables and which functions return them
get_qcew_codebook()

# Total employment — all counties, all years
get_employment(geography = "county")

# Average annual pay — specific counties
get_wage_salary(geoids = c("33009", "33011"), years = 2010:2023)

# Employment by BLS supersector
get_sector_employment(geography = "county", years = 2018:2023,
                      sector_type = "BLS", value_type = "share")

# Employment concentration (HHI)
get_employment_concentration(geography = "county", years = 2023)
```

## Core functions

| Function | Returns | Key parameters |
|---|---|---|
| `get_employment()` | Total average annual employment | `geography`, `geoids`, `years` |
| `get_wage_salary()` | Total average annual pay | `geography`, `geoids`, `years` |
| `get_sector_employment()` | Employment by sector | + `sector_type`, `value_type` |
| `get_sector_wages()` | Average pay by sector | + `sector_type` |
| `get_employment_concentration()` | Employment HHI | `geography`, `geoids`, `years` |

All functions return tidy (long) format:
`geoid | year | sector | variable | value`. Rate and share variables
include an `agg_var` column — the employment weight for computing
correctly weighted averages across geographies.

## Sector classifications

QCEW sector data is available under two classification systems,
controlled by the `sector_type` parameter in `get_sector_employment()`
and `get_sector_wages()`.

### BLS supersectors (`sector_type = "BLS"`)

The default classification follows BLS's own QCEW supersector
groupings — 11 aggregations of 2-digit NAICS industries:

| Sector (variable prefix) | BLS supersector | NAICS codes |
|---|---|---|
| `natural_resources_mining` | Natural Resources and Mining | 11, 21 |
| `construction` | Construction | 23 |
| `manufacturing` | Manufacturing | 31–33 |
| `trade_transportation_utilities` | Trade, Transportation, and Utilities | 22, 42, 44–45, 48–49 |
| `information` | Information | 51 |
| `financial_activities` | Financial Activities | 52–53 |
| `professional_business_services` | Professional and Business Services | 54–56 |
| `education_health_services` | Education and Health Services | 61–62 |
| `leisure_hospitality` | Leisure and Hospitality | 71–72 |
| `other_services` | Other Services | 81 |
| `public_administration` | Public Administration | 92 |

BLS publishes government employment under ownership codes (Federal,
State, Local Government) rather than as a NAICS industry — CORI
derives public administration employment by summing rows where
ownership is Federal, State, or Local Government.

### CORI supersectors (`sector_type = "CORI"`)

CORI aggregates the 11 BLS supersectors into 3 custom groupings based
on whether an industry primarily serves tradable (national/
international) or local demand, loosely following the methodology in
Eckert (2018), *Growing Apart: Tradable Services and the Fragmentation
of the US Economy*.

| CORI supersector | Constituent BLS supersectors | Rationale |
|---|---|---|
| **Tradable Goods** | Natural Resources & Mining, Construction, Manufacturing, Trade, Transportation & Utilities | Industries producing or moving physical goods that cross geographic boundaries and face national or international competition |
| **Tradable Services** | Information, Financial Activities, Professional & Business Services | Knowledge-intensive service industries whose output can be delivered remotely, with markets extending beyond the local area |
| **Local Services** | Education & Health Services, Leisure & Hospitality, Other Services, Public Administration | Industries primarily serving local demand — residents and institutions within the geography |

::: callout-important
## CORI supersector limitations

CORI supersectors are available as **employment shares only**, not
counts — multiply `value` (share) by `agg_var` (total employment) to
approximate counts. This is an opinionated aggregation; use
`sector_type = "BLS"` when the source-level groupings are needed, or
to build a different custom aggregation.
:::

## Geography

| `geography` | `geoid` format | Coverage |
|---|---|---|
| `"county"` | 5-digit FIPS | ~3,100 U.S. counties |
| `"state"` | 2-digit FIPS | 50 states + DC |

```{r}
#| eval: false

# Specific counties
get_employment(geoids = c("33009", "33011"), years = 2010:2023)

# Specific states
get_wage_salary(geoids = c("33", "50", "23"), years = 2015:2023)
```

## Key variables

| Variable | Function | Unit | `agg_var` |
|---|---|---|---|
| `employment` | `get_employment`, `get_sector_employment` | workers | — |
| `avg_pay` | `get_wage_salary`, `get_sector_wages` | nominal $/worker | total employment (wages) or sector employment (sector wages) |
| `emp_share` | `get_sector_employment` | proportion (0–1) | total employment |
| `employment_hhi` | `get_employment_concentration` | index (0–10,000) | — |

Pay values are **nominal dollars** — deflate using a BLS price index
such as the Consumer Price Index (CPI-U) or Employment Cost Index
(ECI) to convert to real dollars.

## Common use cases

### Employment-weighted aggregation

```{r}
#| eval: false

library(dplyr)

pay <- get_wage_salary(geography = "county", years = 2023)

# Employment-weighted average pay across all counties
pay |>
  summarize(
    weighted_avg_pay = sum(value * agg_var, na.rm = TRUE) /
                       sum(agg_var, na.rm = TRUE)
  )
```

### Rural vs. nonrural employment since the Great Recession

```{r}
#| eval: false

library(ruraldefinitions)

rural_definition <- cbsa_2023 |> select(geoid, is_rural)

county_comparison <- get_employment(geography = "county", years = 2007:2023) |>
  left_join(rural_definition, by = "geoid") |>
  filter(!is.na(is_rural)) |>
  group_by(year, is_rural) |>
  summarize(total_employment = sum(value, na.rm = TRUE), .groups = "drop")
```

CORI's own analysis using this pattern found rural employment has been
set back twice since 2007 — a more fragile recovery than nonrural
counties — and that within rural counties, the West led the recovery
by Census region.

### Industry concentration by region

`get_employment_concentration()`'s HHI values are directly comparable
across counties. CORI's own analysis found rural Midwest counties have
more concentrated (less diversified) economies than other rural
regions.

## Vintages

Each processed snapshot is tagged with a vintage (e.g.
`"vintage_2025"`). A `_LATEST` pointer on S3 always resolves to the
most recent vintage, so existing code picks up new data automatically:

```{r}
#| eval: false

# Check current vintage
latest_qcew_vintage()
```

## Related resources

### Datasets

-   [QCEW Employment & Wages](/datasets/qcew-employment-wages/) —
    Source dataset documentation

### Related packages

-   [ruraldefinitions](/packages/ruraldefinitions/) — Rural/nonrural
    county classification used throughout the comparison examples above

### External links

-   [BLS QCEW Homepage](https://www.bls.gov/cew/)
-   [BLS QCEW Industry Supersectors](https://www.bls.gov/cew/classifications/industry/industry-supersectors.htm)
-   [Consumer Price Index (CPI-U)](https://www.bls.gov/cpi/)
-   [Employment Cost Index (ECI)](https://www.bls.gov/eci/)
-   Eckert, F. (2018). [*Growing Apart: Tradable Services and the
    Fragmentation of the US Economy*](https://economics.nd.edu/assets/303413/eckert_jmp_2018.pdf).
    Job Market Paper, University of Notre Dame.

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data.qcew)
-   [Full README](https://github.com/ruralinnovation/cori.data.qcew/blob/main/README.md)
-   [Vignette: Introduction](https://github.com/ruralinnovation/cori.data.qcew/blob/main/vignettes/introduction.Rmd)
-   [Vignette: Rural Employment Trends](https://github.com/ruralinnovation/cori.data.qcew/blob/main/vignettes/rural-employment-trends.Rmd)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.qcew/issues).
