---
title: ruraldefinitions
description: R package aggregating federal rural definitions and making them accessible
  to R programmers. County-level definitions are bundled as package data; tract-level
  definitions are loaded from S3 via get_definition(). Supports CBSA, RUCC, NCHS,
  UIC, RUCA, and CORI rural classifications across multiple vintages.
author: Drew Rosebush
date: '2026-08-31'
categories:
  - R Package
  - Rural Development
tags:
  - Rural
  - CBSA
  - RUCC
  - RUCA
  - Geographic Crosswalk
  - County-Level
  - Rural Definitions
package:
  name: ruraldefinitions
  githubUrl: https://github.com/ruralinnovation/ruraldefinitions
  installCommand: remotes::install_github("ruralinnovation/ruraldefinitions")
  status: development
  version: 0.2.0
  maintainer: Drew Rosebush
  featured: true
usesDatasets:
  - usda-county-typology
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `ruraldefinitions` package aggregates federal rural classification
systems and makes them accessible in R without manual downloads or
crosswalk construction. County-level definitions are bundled directly as
package data — accessible immediately after loading the package. Tract-
and block-level definitions are available via `get_definition()`, which
reads from an S3 bucket.

The package supports seven classification systems across multiple
vintages: the CORI definition, CBSA metropolitan/micropolitan
delineations, USDA Rural-Urban Commuting Area (RUCA) codes, USDA
Rural-Urban Continuum Codes (RUCC), USDA Urban Influence Codes (UIC),
National Center for Health Statistics (NCHS) classification, and the
Census Bureau's block-level rural definition. Every other `cori.data.*`
package uses `ruraldefinitions` to attach rural/nonrural flags to
analysis datasets.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/ruraldefinitions")
```

```{r}
#| eval: false
library(ruraldefinitions)

# County-level definitions — available directly as package data
head(cbsa_2023)       # CBSA 2023 (CORI standard for county analysis)
head(rucc_2023)       # USDA Rural-Urban Continuum Codes 2023
head(nchs_2023)       # NCHS Urban-Rural Classification 2023
head(uic_2024)        # USDA Urban Influence Codes 2024

# Tract-level definitions — loaded from S3
cori_def  <- get_definition("cori", 2020)   # CORI hybrid definition
ruca_def  <- get_definition("ruca", 2020)   # USDA RUCA codes

# Attach rural flag to any county-level dataset
library(dplyr)
my_data |>
  left_join(select(cbsa_2023, geoid, is_rural), by = "geoid")
```

## What This Data Tells Us About Rural

"Rural" is not a single definition. The federal government uses multiple
classification systems — CBSA metropolitan/micropolitan delineations,
USDA Rural-Urban Continuum Codes, NCHS Urban-Rural Classification, USDA
Urban Influence Codes — each designed for different purposes and drawing
the rural/nonrural boundary in different places. A county classified as
rural under one system may not be under another. Researchers who select
a definition without considering what it measures risk drawing
conclusions that reflect the classification rather than the underlying
reality.

CORI's standard for county-level economic analysis is the CBSA-based
definition (OMB metropolitan/micropolitan delineations), which classifies
counties as rural if they fall outside a Core-Based Statistical Area.
This definition is appropriate for most economic analysis because it
reflects labor market integration — counties inside CBSAs have economic
ties to urban cores that counties outside do not. It is not the only
valid choice: USDA's Rural-Urban Continuum Codes offer a gradient rather
than a binary, capturing the range from completely rural to adjacent to
large metro; the NCHS classification is designed specifically for health
research contexts.

The choice of definition is a methodological decision with real
consequences. CORI's research documents patterns that differ between
rural and nonrural counties — in employment composition, broadband
coverage, business formation, and capital access — but the magnitude and
the set of counties involved vary depending on which definition is
applied. Making that choice explicit and defensible is part of credible
rural analysis.

One additional dimension the data makes visible: the misconception that
rural communities are homogeneously white renders people of color
invisible in data, policy, and practice. Rural communities include
significant Black, Hispanic, Native American, and Asian populations whose
experiences and economic assets are not captured in analyses that treat
rural as a monolithic category. Rural classifications are most useful
when paired with demographic data, not used as a substitute for it.

## Available Definitions

County-level definitions are bundled as package data and available
immediately. Tract-level definitions are accessed via `get_definition()`.

| Name | Year(s) | Geographic Unit | Notes |
|---|---|---|---|
| CORI | 2010, 2020 | Tract | RUCA + CBSA hybrid — CORI's standard for tract-level analysis |
| CBSA | 2018, 2020, 2023 | County | OMB metro/micro delineations — CORI's standard for county-level analysis |
| RUCC | 1974–2023 | County | USDA gradient from metro to most rural |
| NCHS | 1990–2023 | County | Designed for health research contexts |
| UIC | 1993–2024 | County | USDA classification by urbanization and metro adjacency |
| RUCA | 1990–2020 | Tract | USDA commuting-based tract classification |
| Census | 2010, 2020 | Block | Census Bureau block-level definition |

## Selecting a Definition

For county-level economic analysis, CORI recommends `cbsa_2023`. If you
have access to tract data and need to capture rural tracts within metro
counties, use the CORI definition via `get_definition("cori", 2020)`.

```{r}
#| eval: false
# CORI standard: county-level rural flag
rural_xwalk <- cbsa_2023 |>
  dplyr::select(geoid, is_rural)

# Tract-level: includes rural tracts inside metro counties
cori_tract <- get_definition("cori", 2020)
```

CORI's published guidance on definition selection:
- [Defining Rural America](https://ruralinnovation.us/blog/defining-rural-america/)
- [Do No Harm Guide](https://www.urban.org/research/publication/do-no-harm-guide-crafting-equitable-data-narratives)
  — Urban Institute writeup on crafting equitable data narratives

## Definition Maps

*Adapted from the [Rural Definitions Maps
vignette](https://github.com/ruralinnovation/ruraldefinitions/blob/main/vignettes/rural-definitions-maps.Rmd)
on GitHub. See the full vignette for maps of all six definitions.*

The following examples show how to map rural definitions using `sf` and
`tigris`. All maps use the same county and state geometries loaded once
and reused.

```{r}
#| eval: false
library(ruraldefinitions)
library(cori.charts)
library(dplyr)
library(ggplot2)
library(sf)
library(tigris)

load_fonts()
options(tigris_use_cache = TRUE)

# Load geometries once
counties_sf <- tigris::counties(cb = TRUE, year = 2023) |>
  tigris::shift_geometry() |>
  filter(GEOID <= "57") |>
  sf::st_transform(crs = 5070)

states_sf <- tigris::states(cb = TRUE, year = 2023) |>
  tigris::shift_geometry() |>
  filter(GEOID <= "57") |>
  sf::st_transform(crs = 5070)

bbox <- sf::st_bbox(states_sf)
```

### CBSA 2023 — CORI's county standard

```{r}
#| eval: false
colors_cbsa <- c(
  "Metro"    = cori_colors[["Dark Purple"]],
  "Micro"    = "#086656",
  "Non-CBSA" = cori_colors[["Emerald"]]
)

map_cbsa <- counties_sf |>
  left_join(cbsa_2023, by = c("GEOID" = "geoid")) |>
  mutate(rural_def = factor(rural_def, levels = names(colors_cbsa))) |>
  group_by(rural_def) |>
  summarise(geometry = sf::st_union(geometry), .groups = "drop")

ggplot() +
  geom_sf(data = map_cbsa, aes(fill = rural_def), color = NA) +
  geom_sf(data = states_sf, fill = NA, color = "black", linewidth = 0.3) +
  scale_fill_manual(name = NULL, values = colors_cbsa, na.value = "#ffffff") +
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]),
           ylim = c(bbox["ymin"], bbox["ymax"]), expand = FALSE) +
  labs(
    title   = "Core-Based Statistical Areas (CBSA), 2023",
    subtitle = "County-level OMB metropolitan/micropolitan delineation",
    caption = "Source: U.S. Office of Management and Budget, 2023 delineations."
  ) +
  theme_cori_map() +
  theme(legend.position = "top", legend.direction = "horizontal")
```

### CORI 2020 — tract-level hybrid definition

```{r}
#| eval: false
state_fips <- tigris::fips_codes |>
  dplyr::distinct(state_code) |>
  dplyr::filter(state_code <= "56") |>
  dplyr::pull(state_code)

tracts_sf <- purrr::map_dfr(
  state_fips,
  ~tigris::tracts(state = .x, cb = TRUE, year = 2020, progress_bar = FALSE)
) |>
  tigris::shift_geometry() |>
  sf::st_transform(crs = 5070)

cori_raw <- get_definition("cori", 2020)

colors_cori <- c(
  "Metro"                      = cori_colors[["Dark Purple"]],
  "Nonmetro/RUCA 4+ carveouts" = cori_colors[["Emerald"]]
)

map_cori <- tracts_sf |>
  left_join(cori_raw, by = c("GEOID" = "geoid")) |>
  filter(!is.na(rural_def)) |>
  mutate(
    display = case_when(
      rural_def == "Metro county" ~ "Metro",
      rural_def %in% c("Nonmetro county",
                       "Rural tract in metro county") ~ "Nonmetro/RUCA 4+ carveouts"
    ),
    display = factor(display, levels = names(colors_cori))
  ) |>
  group_by(display) |>
  summarise(geometry = sf::st_union(geometry), .groups = "drop")

ggplot() +
  geom_sf(data = map_cori, aes(fill = display), color = NA) +
  geom_sf(data = states_sf, fill = NA, color = "black", linewidth = 0.3) +
  scale_fill_manual(name = NULL, values = colors_cori, na.value = "#ffffff") +
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]),
           ylim = c(bbox["ymin"], bbox["ymax"]), expand = FALSE) +
  labs(
    title    = "CORI Rural Definition, 2020",
    subtitle = "Nonmetro counties (CBSA 2023) plus rural tracts (RUCA 4+) in metro counties",
    caption  = "Source: Center on Rural Innovation."
  ) +
  theme_cori_map() +
  theme(legend.position = "top", legend.direction = "horizontal")
```

## Related

-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `ruraldefinitions`
-   **Dataset:** [USDA County Typology](/datasets/usda-county-typology/)
    — further characterizes rural counties by economic and demographic
    type
