---
title: "ruraldefinitions"
description: "R package for accessing federal rural definitions at multiple geographic levels"
author: "Drew Rosebush"
date: "2026-08-26"
categories: ["R Package", "Rural Classification"]
tags: ["USDA", "Census", "RUCA", "Rural Definitions", "Geography"]
package:
  name: "ruraldefinitions"
  githubUrl: "https://github.com/ruralinnovation/ruraldefinitions"
  installCommand: 'remotes::install_github("ruralinnovation/ruraldefinitions")'
  status: "stable"
  version: "1.0.0"
  maintainer: "Drew Rosebush"
  featured: true
usesDatasets: []
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `ruraldefinitions` package aggregates seven federal rural
classification systems and makes them accessible to R programmers.
County- and block-level definitions are loaded directly into the
package as data; tract-level definitions (RUCA, CORI) are downloaded
from S3 on demand via the package's single exported function,
`get_definition()`.

::: callout-note
## What's in the package

-   **One-stop access** — seven major federal rural definitions in one
    package
-   **Standardized format** — every definition shares the same
    columns: `geoid`, `name`, `year`, `rural_def`, `is_rural`
-   **Historical depth** — vintages spanning 1974–2024 across the
    seven systems
-   **Three geographic levels** — county, tract, and block
-   **Expert guidance** — built-in recommendations for which
    definition to use, by geography level
:::

## Installation

::: panel-tabset
## GitHub

```{r}
#| eval: false

remotes::install_github("ruralinnovation/ruraldefinitions")
```

## AWS S3 access

For non-R users, CSV files for every definition and vintage are
available directly from S3 — see the table below for per-file links.
:::

## Quick start

```{r}
#| eval: false

library(ruraldefinitions)

# County- and block-level definitions are pre-loaded package data
head(cbsa_2023)
head(rucc_2023)

# Tract-level definitions download from S3 on first use
cori_2020 <- get_definition("cori", 2020)
ruca_2010 <- get_definition("ruca", 2010)

# Every definition shares the same columns:
# geoid | name | year | rural_def | is_rural ("Rural" / "Nonrural")
```

## Available definitions

| Name | Year(s) | Geography | Access |
|---|---|---|---|
| **Census** | 2010, 2020 | Block | Pre-loaded package data |
| **RUCA** | 1990, 2000, 2010, 2020 | Tract | `get_definition()`, downloads from S3 |
| **CORI** | 2010, 2020 | Tract | `get_definition()`, downloads from S3 |
| **CBSA** | 2018, 2020, 2023 | County | Pre-loaded package data |
| **NCHS** | 1990, 2006, 2013, 2023 | County | Pre-loaded package data |
| **RUCC** | 1974, 1983, 1993, 2003, 2013, 2023 | County | Pre-loaded package data |
| **UIC** | 1993, 2003, 2013, 2024 | County | Pre-loaded package data |

## Selecting a definition

::: callout-tip
## Choosing the right definition

**For county-level analysis:** use the **CBSA** ("Nonmetro") definition
— it aligns with OMB standards and is best for state/regional
comparisons.

**For tract-level analysis:** use the **CORI** definition — it
captures rural tracts within otherwise-metropolitan counties, which
county-level definitions alone miss.

Read more in CORI's piece on [defining rural
America](https://ruralinnovation.us/blog/defining-rural-america/) or
the Urban Institute's [Do No Harm
Guide](https://www.urban.org/research/publication/do-no-harm-guide-crafting-equitable-data-narratives)
on crafting equitable data narratives.
:::

## Definition details

**CORI** classifies Census tracts as rural if they sit within a
nonmetropolitan county (OMB CBSA definition) or score 4 or higher on
USDA's Rural-Urban Commuting Area (RUCA) codes. Two vintages are
available: CORI 2010 (RUCA 2010 + CBSA 2018) and CORI 2020 (RUCA 2020 +
CBSA 2023), each anchored to its own Census boundaries.

**RUCA (Rural-Urban Commuting Area Codes)** classifies Census tracts
by economic integration — population density, urbanization, and
commuting patterns. Published by USDA. RUCA 2010 and 2020 are **not**
directly comparable, due to changes in the Census Bureau's urban-area
delineation criteria and updates to the urban-core classification
threshold between vintages.

**CBSA (Core-Based Statistical Areas)** designates counties as
metropolitan, micropolitan, or noncore by population size and the
presence of urban centers; micropolitan and noncore counties count as
nonmetropolitan. Defined by OMB, updated roughly decennially with
periodic mid-decade revisions — the 2018 vintage is one such
mid-decade update.

**NCHS (National Center for Health Statistics)** subdivides the OMB
definition into four metropolitan and two nonmetropolitan categories;
metropolitan counties of 1 million+ population are further split into
central (e.g. inner cities) and fringe (e.g. suburbs).

**RUCC (Rural-Urban Continuum Codes)** distinguishes metropolitan
counties by population size and nonmetropolitan counties by degree of
urbanization and proximity to a metro area. Published by USDA.

**UIC (Urban Influence Codes)** classifies counties by population
size, degree of urbanization, and adjacency to metropolitan areas.
Published by USDA.

**Census** defines rural as Census blocks not located within an urban
area, using population-size and housing-density thresholds.

## Common use cases

### Basic rural classification

```{r}
#| eval: false

library(dplyr)

rural_tracts <- get_definition("cori", 2020)

rural_only <- rural_tracts |> filter(is_rural == "Rural")

rural_tracts |> count(rural_def, is_rural)
```

### Comparing definitions

```{r}
#| eval: false

ruca_2010 <- get_definition("ruca", 2010)
cori_2020 <- get_definition("cori", 2020)

comparison <- ruca_2010 |>
  left_join(cori_2020, by = "geoid", suffix = c("_ruca", "_cori")) |>
  count(is_rural_ruca, is_rural_cori)
```

### Historical analysis

```{r}
#| eval: false

# Pre-loaded data across vintages
historical <- rucc_2023 |>
  select(geoid, rural_def_2023 = rural_def) |>
  left_join(
    rucc_2013 |> select(geoid, rural_def_2013 = rural_def),
    by = "geoid"
  )

# Counties that changed classification
changed <- historical |> filter(rural_def_2023 != rural_def_2013)
```

### Mapping a definition

```{r}
#| eval: false

library(sf)
library(tigris)
library(ggplot2)

counties <- counties(cb = TRUE, resolution = "20m") |>
  mutate(geoid = GEOID)

counties |>
  left_join(cbsa_2023, by = "geoid") |>
  ggplot(aes(fill = is_rural)) +
  geom_sf(linewidth = 0.05) +
  labs(title = "Core-Based Statistical Areas (CBSA), 2023")
```

## Data structure

Every definition shares the same structure:

```
geoid       - Geographic identifier (FIPS code)
name        - Definition name (e.g., "CORI", "RUCA")
year        - Year of classification
rural_def   - Detailed classification (e.g., "Metropolitan", "Micropolitan")
is_rural    - Binary: "Rural" or "Nonrural"
```

## Performance considerations

::: callout-tip
## Loading strategy

-   **County and block data** — pre-loaded, instant access
-   **Tract data** — downloads from S3 (~1–2 MB) on first use; cache
    locally to avoid repeated downloads
:::

## Related resources

### External links

-   [USDA ERS Rural Classifications](https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/)
-   [Census Urban/Rural Classification](https://www.census.gov/programs-surveys/geography/guidance/geo-areas/urban-rural.html)
-   [RUCA Data](https://www.ers.usda.gov/data-products/rural-urban-commuting-area-codes/)
-   [RUCC Data](https://www.ers.usda.gov/data-products/rural-urban-continuum-codes/)
-   [UIC Data](https://www.ers.usda.gov/data-products/urban-influence-codes/)
-   [NCHS Urban-Rural Classification](https://www.cdc.gov/nchs/data-analysis-tools/urban-rural.html)
-   [CORI: Defining Rural America](https://ruralinnovation.us/blog/defining-rural-america/)
-   [Urban Institute: Do No Harm Guide](https://www.urban.org/research/publication/do-no-harm-guide-crafting-equitable-data-narratives)

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/ruraldefinitions)
-   [Full README](https://github.com/ruralinnovation/ruraldefinitions/blob/main/README.md)
-   [Vignette: Rural Definitions Maps](https://github.com/ruralinnovation/ruraldefinitions/blob/main/vignettes/rural-definitions-maps.Rmd)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/ruraldefinitions/issues).
