---
title: "cori.data"
description: "CORI meta-package that installs and attaches all cori.data.* companion packages in a single call"
author: "John Hall"
date: "2026-08-26"
categories: ["R Package", "Data Infrastructure"]
tags: ["Meta-Package", "Rural Data", "Tidy Data", "Ecosystem"]
package:
  name: "cori.data"
  githubUrl: "https://github.com/ruralinnovation/cori.data"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data")'
  status: "stable"
  version: "2.0.0"
  maintainer: "John Hall"
  featured: true
usesDatasets: []
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

> Rural data, made accessible.

`cori.data` is the meta-package for the `cori.data.*` family. It has
no functions of its own — installing it pulls in every companion
package via its `Depends:`/`Remotes:` declarations; loading it attaches
them all, so `get_employment()`, `get_population()`,
`get_definition()`, and dozens more become available without separate
`library()` calls for each one.

[`cori.data.s3`](/packages/cori-data-s3/) is the shared S3/cloud-access
layer underneath the family — credential management, a DuckDB S3
connection helper, and S3 object functions every companion package
builds on. Reads require no local AWS setup; the package falls back to
temporary, read-only credentials from CORI's credential-vending
endpoint automatically.

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data")
```

Because `cori.data` declares every companion package under
`Depends:`/`Remotes:`, `remotes::install_github()` resolves and
installs the full family in this one call — no separate installer
function is needed.

## The ecosystem

| Package | What it measures | Geography | Years |
|---|---|---|---|
| [cori.data.bds](/packages/cori-data-bds/) | Business dynamics (Census BDS) | county, state, nation | 1978–present |
| [cori.data.bfs](/packages/cori-data-bfs/) | Business applications (Census BFS) | county, state, nation | 2005–present |
| [cori.data.bps](/packages/cori-data-bps/) | Building permits (Census BPS) | county, state, nation | 2000–present |
| [cori.data.hu](/packages/cori-data-hu/) | Housing units (Census HU) | county, state, nation | 2000–present |
| [cori.data.fcc](/packages/cori-data-fcc/) | Broadband coverage (FCC NBM) | county | 2022–present |
| [cori.data.pep](/packages/cori-data-pep/) | Population (Census PEP) | county, state, nation | 2000–present |
| [cori.data.qcew](/packages/cori-data-qcew/) | Employment & wages (BLS QCEW) | county, state | 1990–present |
| [ruraldefinitions](/packages/ruraldefinitions/) | Rural classifications (OMB, USDA ERS, NCHS) | county | multiple vintages |

## A consistent interface

Every data package speaks the same language. Once you learn one, you
know them all:

```{r}
#| eval: false

library(cori.data.qcew)
library(cori.data.pep)
library(cori.data.bds)

# Employment by county, last 5 years
get_employment(geography = "county", years = 2019:2023)

# Population by county, last 5 years
get_population(geography = "county", years = 2019:2023)

# Business dynamics by county, last 5 years
get_business_dynamics(geography = "county", years = 2019:2023)

# Or filter to specific places with FIPS codes
get_employment(geoids = c("33009", "54011", "30001"), years = 2019:2023)
```

All functions return tidy (long) format data — one row per
`geoid × year × variable`, ready for `dplyr`, `ggplot2`, and `tidyr`.

## Rural classifications

`ruraldefinitions` provides rural classifications across multiple
vintages and methodologies. County- and block-level definitions are
loaded as package data; tract-level definitions (RUCA, CORI) are
downloaded from S3 on demand via `get_definition()`:

```{r}
#| eval: false

library(ruraldefinitions)

# CBSA-based rural/nonrural (the CORI standard for county analysis) —
# pre-loaded package data, not a function call
rural <- cbsa_2023
# Returns: geoid | name | year | rural_def | is_rural ("Rural" / "Nonrural")

# Other pre-loaded county-level definitions
rucc_2023  # USDA Rural-Urban Continuum Codes
nchs_2023  # NCHS Urban-Rural Classification Scheme
uic_2024   # USDA Urban Influence Codes

# Tract-level definitions download from S3 on first use
cori_2020 <- get_definition("cori", 2020)
```

## Cross-package analysis

The real power of the ecosystem comes from combining packages — an
employment rate built from two different sources and stratified by
rural status:

```{r}
#| eval: false

library(cori.data.qcew)
library(cori.data.pep)
library(ruraldefinitions)
library(dplyr)

# Employment rate = covered employment / working-age population (16+)
emp <- get_employment(geography = "county", years = 2010:2023)
pop <- get_population(geography = "county", years = 2010:2023,
                      variables = "population_16plus")

emp_rate <- emp |>
  filter(variable == "employment") |>
  select(geoid, year, employment = value) |>
  left_join(
    pop |> filter(variable == "population_16plus") |>
      select(geoid, year, pop_16plus = value),
    by = c("geoid", "year")
  ) |>
  mutate(emp_rate = employment / pop_16plus)

# Add rural classification and summarize
emp_summary <- emp_rate |>
  left_join(cbsa_2023 |> select(geoid, is_rural), by = "geoid") |>
  filter(!is.na(is_rural)) |>
  group_by(year, is_rural) |>
  summarize(
    avg_emp_rate = weighted.mean(emp_rate, employment, na.rm = TRUE),
    .groups = "drop"
  )
```

## Design principles

All packages in the family follow the same design principles: tidy
output, consistent parameters (`geography`, `geoids`, `years`,
`variables`), plain-English variable names, and no infrastructure
knowledge required to get data.

## Terms of use

Data in this ecosystem are curated from U.S. federal agency sources and
provided "as is" without warranty of any kind. CORI processes and
packages these data to improve accessibility for rural research but
cannot guarantee their accuracy, completeness, or timeliness, and
cannot vouch for analyses derived from these data after retrieval.

The following notices are required by the underlying data providers'
terms of service:

-   **Census Bureau:** This product uses the Census Bureau Data API but
    is not endorsed or certified by the Census Bureau.
-   **BLS:** BLS.gov cannot vouch for the data or analyses derived from
    these data after the data have been retrieved from BLS.gov.
-   **FCC:** Broadband availability data from the FCC Broadband Data
    Collection are in the public domain (17 U.S.C. § 105).

Full terms of service:

-   [Census Bureau API Terms of Service](https://www.census.gov/data/developers/about/terms-of-service.html)
-   [BLS Terms of Service](https://www.bls.gov/developers/termsOfService.htm)
-   [FCC National Broadband Map — License and Attribution](https://broadbandmap.fcc.gov/about)

## Contributing

CORI is a small team building tools for rural researchers and welcomes
collaborators. Open an issue on the
[`cori.data` GitHub repo](https://github.com/ruralinnovation/cori.data)
or any of the companion package repositories for bugs, features, or
data the ecosystem should add.

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data)
-   [Full README](https://github.com/ruralinnovation/cori.data/blob/main/README.md)
-   [Getting Started Vignette](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/getting-started.Rmd)
-   [Vignette: Tidy Data](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/tidy-data.Rmd)
-   [Vignette: Weighted Averages](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/weighted-averages.Rmd)
-   [Vignette: Sector Analysis](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/sector-analysis.Rmd)
-   [Vignette: Employment Rate](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/employment-rate.Rmd)
-   [Vignette: Broadband](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/broadband.Rmd)

---

**Contact:** John Hall (john.hall@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data/issues).
