---
title: "cori.data"
description: "CORI's rural data ecosystem — a family of R packages providing unified access to employment, population, business, housing, and broadband data across rural America"
author: "Mapping & Data Analytics Team"
date: "2026-08-27"
categories: ["R Package", "Data Infrastructure"]
tags: ["Meta-Package", "Rural Data", "Tidy Data", "Ecosystem"]
package:
  name: "cori.data"
  githubUrl: "https://github.com/ruralinnovation/cori.data"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data")'
  status: "stable"
  version: "2.0.0"
  maintainer: "John Hall"
  featured: true
usesDatasets: []
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

`cori.data` is the meta-package for CORI's rural data ecosystem — a
growing collection of R packages providing unified access to employment,
population, business, housing, and broadband data across rural America.
Installing it pulls in every companion package; loading it attaches them
all, so `get_employment()`, `get_population()`, `get_definition()`, and
more are available without separate `library()` calls.

Not all federal data sources are designed with rural communities in mind.
CORI's data team has done the methodological work — identifying the right
sources, applying consistent rural classifications, and packaging data in
a format that works across rural geographies — so analysts can focus on
the research question, not the data infrastructure.

## Installation & Quick Start

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data")
```

```{r}
#| eval: false

library(cori.data)

# Employment, population, and business dynamics —
# all joinable on geoid + year, ready in minutes
get_employment(geography = "county", years = 2019:2023)
get_population(geography = "county", years = 2019:2023)
get_business_dynamics(geography = "county", years = 2019:2023)
```

## The Ecosystem

| Package | What it measures | Geography | Years |
|---|---|---|---|
| [cori.data.qcew](/packages/cori-data-qcew/) | Employment & wages (BLS QCEW) | county, state | 1990–present |
| [cori.data.bds](/packages/cori-data-bds/) | Business dynamics (Census BDS) | county, state, nation | 1978–present |
| [cori.data.pep](/packages/cori-data-pep/) | Population (Census PEP) | county, state, nation | 2000–present |
| [cori.data.bfs](/packages/cori-data-bfs/) | Business applications (Census BFS) | county, state, nation | 2004–present |
| [cori.data.bps](/packages/cori-data-bps/) | Building permits (Census BPS) | county, state, nation | 2000–present |
| [cori.data.hu](/packages/cori-data-hu/) | Housing units (Census HU) | county, state, nation | 2000–present |
| [cori.data.fcc](/packages/cori-data-fcc/) | Broadband coverage (FCC NBM) | county | 2022–present |
| [ruraldefinitions](/packages/ruraldefinitions/) | Rural classifications (OMB, USDA ERS, NCHS) | county | multiple vintages |

## What This Data Tells Us About Rural

Rural economies are diverse, complex, and deeply shaped by the assets
and advantages they have accumulated over time — natural resources,
agricultural land, industrial infrastructure, geographic position,
community institutions, and workforce traditions. Some rural communities
are thriving, leveraging those assets in new ways as remote work,
entrepreneurship, and targeted investment open new pathways. Others are
navigating structural transitions as legacy industries shift. What they
share is a foundation of real assets whose value in the modern economy
depends on the conditions surrounding them.

Understanding which assets a community holds, how those assets connect
to current economic conditions, and what they might support going
forward requires data that spans multiple dimensions simultaneously.
Employment trends read alongside business formation rates, population
dynamics, broadband coverage, and sector composition tell a richer
story than any single source alone. Across CORI's research, communities
with strong asset bases across multiple dimensions tend to show more
resilient economic trajectories than those concentrated in a single
advantage — and human capital, quality of life, and the capacity to
retain and attract working-age residents consistently emerge as
important complements to physical infrastructure.

The ecosystem exists because these questions require data that is
consistent, joinable, and contextualized for rural analysis — and
because the methodological decisions that make cross-source rural
analysis tractable (which rural definition to apply, how to weight
aggregations across counties of very different sizes, which sector
taxonomy to use) are embedded in the packages rather than left to each
researcher to reinvent.

## Getting Started

*Adapted from the [Getting Started
vignette](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/getting-started.Rmd)
on GitHub.*

### A consistent interface

Every package uses the same four parameters. Learn the pattern once and
it works everywhere:

```{r}
#| eval: false

library(cori.data.qcew)
library(cori.data.pep)
library(cori.data.bds)

# geography = "county", "state", or "nation"
# geoids    = character vector of FIPS codes (overrides geography)
# years     = integer vector of years
# variables = which variables to return (NULL = all)

get_employment(geography = "county", years = 2019:2023)
get_population(geography = "county", years = 2019:2023)
get_business_dynamics(geography = "county", years = 2019:2023)

# Filter to specific places with FIPS codes
get_employment(geoids = c("33009", "54011", "30001"), years = 2019:2023)
```

All functions return **tidy (long) format** data — one row per
`geoid × year × variable`:

```
geoid   year  variable      value
33009   2023  employment    28450
33009   2022  employment    27810
54011   2023  employment    12340
```

### Your first pull

```{r}
#| eval: false

library(cori.data.qcew)
library(dplyr)

emp <- get_employment(geography = "county", years = 2019:2023)

glimpse(emp)

# Which counties had the most employment in 2023?
emp |>
  filter(year == 2023) |>
  slice_max(value, n = 10)
```

### Rural classifications

The `ruraldefinitions` package provides county-level rural/nonrural
classifications across several methodologies. CORI's standard for
county analysis is the CBSA-based definition:

```{r}
#| eval: false

library(ruraldefinitions)

# CBSA-based (the CORI standard) — pre-loaded, no function call needed
rural <- cbsa_2023
# Returns: geoid | name | year | rural_def | is_rural ("Rural" / "Nonrural")

# Other available county-level definitions
rucc_2023  # USDA Rural-Urban Continuum Codes
nchs_2023  # NCHS Urban-Rural Classification Scheme
uic_2024   # USDA Urban Influence Codes
```

## Working with Tidy Data

*Adapted from the [Working with Tidy
Data](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/tidy-data.Rmd)
vignette on GitHub.*

The most powerful analyses come from combining packages — employment
with population, business formation with building permits, broadband
with wages. What makes that combination frictionless is a shared data
contract: every function in every package returns the same structure,
with the same column names, at the same geographic identifiers. Learn
these patterns once and they work across the entire ecosystem.

### Filtering by variable

```{r}
#| eval: false

bd <- get_business_dynamics(geography = "county", years = 2018:2023)

# Single variable
estabs <- bd |> filter(variable == "establishments")

# Multiple variables
rates <- bd |> filter(variable %in% c("estab_entry_rate", "estab_exit_rate"))
```

### Pivoting to wide format

```{r}
#| eval: false

bd_wide <- get_business_dynamics(geography = "county", years = 2022:2023) |>
  pivot_wider(names_from = variable, values_from = value)

# Now: geoid | year | establishments | estab_entries | estab_entry_rate | ...
```

### Joining across packages

Because every package returns the same `geoid + year` identifiers,
joining across packages is a simple `left_join()`:

```{r}
#| eval: false

# Employment + population + business dynamics in one county profile
emp <- get_employment(geography = "county", years = 2023) |>
  filter(variable == "employment") |>
  select(geoid, employment = value)

pop <- get_population(geography = "county", years = 2023) |>
  filter(variable == "population") |>
  select(geoid, population = value)

estab <- get_business_dynamics(geography = "county", years = 2023) |>
  filter(variable == "establishments") |>
  select(geoid, establishments = value)

county_profile <- emp |>
  left_join(pop,   by = "geoid") |>
  left_join(estab, by = "geoid") |>
  mutate(
    emp_per_capita   = employment    / population,
    estab_per_capita = establishments / population
  )
```

### Summarizing by rural status

With a rural classification joined in, aggregating to rural/nonrural
summaries is straightforward — and produces comparisons that are
meaningful for rural research:

```{r}
#| eval: false

library(ruraldefinitions)

rural <- cbsa_2023 |> select(geoid, is_rural)

emp_summary <- get_employment(geography = "county", years = 2018:2023) |>
  filter(variable == "employment") |>
  left_join(rural, by = "geoid") |>
  filter(!is.na(is_rural)) |>
  group_by(year, is_rural) |>
  summarize(
    total_employment = sum(value, na.rm = TRUE),
    n_counties       = n_distinct(geoid),
    .groups = "drop"
  )
```

### Visualizing with cori.charts

Tidy data maps directly to `ggplot2` aesthetics — and `cori.charts`
provides CORI-themed formatting:

```{r}
#| eval: false

library(ggplot2)
library(cori.charts)
load_fonts()

get_business_dynamics(geography = "nation", years = 2000:2023) |>
  filter(variable %in% c("estab_entry_rate", "estab_exit_rate")) |>
  ggplot(aes(x = year, y = value, color = variable)) +
  geom_line(linewidth = 1.5) +
  scale_color_cori(
    palette = "ctg2buor",
    labels  = c("estab_entry_rate" = "Entry rate",
                "estab_exit_rate"  = "Exit rate")
  ) +
  scale_y_continuous(
    labels = scales::label_percent(accuracy = 0.1, scale = 1)
  ) +
  theme_cori() +
  labs(
    title   = "National establishment entry and exit rates, 2000–2023",
    x       = NULL,
    y       = NULL,
    color   = NULL,
    caption = "Source: CORI analysis of Census Bureau Business Dynamics Statistics"
  )
```

## Related

**Packages in this ecosystem:**
- [cori.data.qcew](/packages/cori-data-qcew/) — Employment & wages
- [cori.data.bds](/packages/cori-data-bds/) — Business dynamics
- [cori.data.pep](/packages/cori-data-pep/) — Population estimates
- [cori.data.fcc](/packages/cori-data-fcc/) — Broadband coverage
- [ruraldefinitions](/packages/ruraldefinitions/) — Rural classifications


**Further reading:**
- [Employment Rate vignette](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/employment-rate.Rmd) — cross-package analysis combining QCEW + PEP
- [Sector Analysis vignette](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/sector-analysis.Rmd) — CORI super-sectors and rural economic composition
- [Weighted Averages vignette](https://github.com/ruralinnovation/cori.data/blob/main/vignettes/weighted-averages.Rmd) — why `agg_var` matters

::: callout-note
Data in this ecosystem are curated from U.S. federal agency sources.
This product uses the Census Bureau Data API but is not endorsed or
certified by the Census Bureau. BLS.gov cannot vouch for data or
analyses derived from these data after retrieval from BLS.gov.
:::
