---
title: "cori.data.hu"
description: "Access and analyze U.S. Census Bureau Housing Unit Estimates at the national, state, and county level from 2020 to present"
author: "Drew Rosebush"
date: "2026-05-12"
categories: ["R Package", "Housing"]
tags: ["Housing Units", "Census", "Housing Stock", "County-Level", "S3", "Tidy Data"]
package:
  name: "cori.data.hu"
  githubUrl: "https://github.com/ruralinnovation/cori.data.hu"
  installCommand: 'devtools::install_github("ruralinnovation/cori.data.hu")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["census-housing-units"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.hu` package provides access to the U.S. Census Bureau's
**National, State, and County Housing Unit Totals** — annual estimates
of the total number of housing units in existence, published under the
Population Estimates Program. This package measures housing **stock**
(how many units exist at a point in time), as distinct from construction
activity or permitting flow.

::: callout-note
## Why Use This Package

-   **Total housing stock** — Measures all existing units, not just new
    construction; pairs with `cori.data.bps` for net-new activity
-   **S3-backed access** — Pre-processed data queryable via DuckDB
    without downloading per-state source files
-   **Long history** — Vintage series back to 1970, anchored to each
    decennial census
-   **Methodology-aware** — README documents the components-of-change
    model, vintage gap analysis, and known limitations for longitudinal
    use
:::

## Installation

```{r}
#| eval: false

devtools::install_github("ruralinnovation/cori.data.hu")
```

## Quick Start

```{r}
#| eval: false

library(cori.data.hu)

# Read housing unit estimates for all counties, latest vintage
df <- read_hu_from_s3()

# Variable documentation
get_hu_codebook()

# Most recent available vintage year
latest_hu_vintage()
```

## Key Features

### S3-Backed Data Access

```{r}
#| eval: false

# Read all counties, current vintage
read_hu_from_s3()
```

### Components-of-Change Methodology

The Census Bureau builds housing unit estimates using:

```
2020 Census Housing Units
  + New Residential Construction  (permitted × completion rate + non-permitted)
  + New Mobile Homes              (Manufactured Homes Survey)
  - Housing Units Lost            (AHS-based loss rates + FEMA disaster losses)
= July 1 Annual Housing Unit Estimate
```

## Core Functions

| Function | Category | Description |
|---|---|---|
| `read_hu_from_s3()` | Data Access | Read processed housing unit estimates from S3 via DuckDB |
| `get_hu_codebook()` | Utility | Return variable definitions and metadata |
| `latest_hu_vintage()` | Utility | Return the most recent available vintage year |

## Data Coverage

| Dimension | Coverage |
|---|---|
| **Geography** | National, State, County (5-digit FIPS) |
| **Time Period** | 2020–present (current series); historical back to 1970 |
| **Vintage** | Annual (typically released each May) |
| **Update Frequency** | Annual |
| **Source** | Census Bureau Population Estimates Program |

## Known Limitations

::: callout-important
## Cross-Decade Comparisons

Do **not** compute net change across the 2009→2010 or 2019→2020
boundaries. Each decade series is anchored to its own decennial census
baseline. Subtracting across these resets produces a methodological
artifact, not a real change in housing stock. For year-over-year change
analysis, restrict to a single decade series or use `cori.data.bps`
permit data as a proxy for net new construction.
:::

Additional limitations:

-   **No structure type breakdown** in the public release — pair with
    BPS data for single-family vs. multifamily distinctions
-   **No vacancy or occupancy status** — pair with ACS for vacancy rates
-   **Loss rates are static** — derived from 2009/2011 AHS, unchanged
    across all vintages; demolition patterns may be underestimated in
    active redevelopment markets
-   **Non-permitted construction not distributed to the West** —
    methodology gap for Western counties without permit offices

## Use Cases

-   **Housing supply analysis** — Track total stock growth or decline at
    county level
-   **Rural housing research** — Compare housing stock trends between
    rural and nonrural counties
-   **Per-capita housing ratios** — Combine with PEP population
    estimates for units-per-household analysis
-   **Longitudinal panels** — Build within-decade housing stock panels
    (restrict to single vintage series)

## Related Resources

### Datasets

-   Census Housing Unit Estimates *(slug: `census-housing-units` — node
    pending)* — Source dataset documentation

### Related Packages

-   [cori.data.bds](/packages/cori-data-bds/) — Business Dynamics
    Statistics for economic context
-   [cori.data.pep](/packages/cori-data-pep/) — Population estimates
    for per-capita denominators

### External Links

-   [Census Housing Unit Estimates Dataset Page](https://www.census.gov/data/datasets/time-series/demo/popest/2020s-total-housing-units.html)
-   [Vintage 2024 Methodology (PDF)](https://www2.census.gov/programs-surveys/popest/technical-documentation/methodology/2020-2024/2024-hu-meth.pdf)

::: callout-note
## Dangling references

The following slugs are referenced by this package but do not yet have
nodes in Dataverse. They are intentionally preserved as future content
needs:

-   `dataset/census-housing-units`
:::

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.hu/issues).
