---
title: cori.data.hu
description: Access and analyze U.S. Census Bureau Housing Unit Estimates at the
  national, state, and county level from 2000 to present. Measures total housing
  stock — occupied and vacant — as part of CORI's housing supply data ecosystem.
author: Drew Rosebush
date: '2026-08-31'
categories:
  - R Package
  - Housing
tags:
  - Housing Units
  - Census
  - Housing Stock
  - County-Level
  - S3
  - Tidy Data
  - Rural
package:
  name: cori.data.hu
  githubUrl: https://github.com/ruralinnovation/cori.data.hu
  installCommand: remotes::install_github("ruralinnovation/cori.data.hu")
  status: development
  version: 0.1.0
  maintainer: Drew Rosebush
  featured: false
usesDatasets: []
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.hu` package provides access to U.S. Census Bureau
Housing Unit Estimates at the county, state, and national level from
2000 to the present. It measures the total housing stock — occupied and
vacant units combined — built from the decennial Census baseline plus
annual additions from permitted construction, non-permitted
construction, mobile home shipments, and estimated housing loss.

Housing units data differs from building permits data in an important
way: permits measure new supply being added in a given year, while
housing units measure the total stock available. Combined, the two
datasets support analysis of how supply is changing relative to the
existing base. All data are returned in the standard `cori.data` long format: one row
per `geoid / year / variable`.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/cori.data.hu")
```

```{r}
#| eval: false
library(cori.data.hu)

# All available variables
get_hu_codebook()

# Total housing units — all counties, all years
get_housing_units(geography = "county")

# Housing units per 1,000 population for New Hampshire counties
get_housing_units(
  variables = "housing_units_per_1k_people",
  geoids    = c("33009", "33011"),
  years     = 2010:2024
)
```

## What This Data Tells Us About Rural

Housing is the physical substrate of rural economic development — the
asset that determines whether a community can translate economic momentum
into population and workforce growth. A community may have strong
broadband, an active entrepreneurial ecosystem, employment anchors, and
quality of life assets that attract interest from remote workers and
returnees — and still be unable to grow if there is nowhere for people
to live. In this sense, housing stock is not background infrastructure;
it is the capacity of a place to receive and retain people.

## Introduction

*Adapted from the [Introduction
vignette](https://github.com/ruralinnovation/cori.data.hu/blob/main/vignettes/introduction.Rmd)
on GitHub.*

`cori.data.hu` provides annual housing unit estimates from the U.S.
Census Bureau's National, State, and County Housing Unit Totals,
Data cover 2000–present
and measure the total housing stock built from the decennial Census
baseline plus annual construction and loss estimates.

**Source:** U.S. Census Bureau, Population Estimates Program
**Coverage:** 2000–present, updated annually (release typically May)
**Geography:** County (5-digit FIPS), state (2-digit FIPS), national
(`"00"`)

### Variables

```{r}
#| eval: false
get_hu_codebook() |>
  dplyr::select(variable, label, unit, notes) |>
  knitr::kable()
```

### Reading data

All data are returned in long format: one row per `geoid / year /
variable`.

```{r}
#| eval: false
df <- get_housing_units(geography = "county")
dplyr::glimpse(df)
```

Filter to specific variables, years, or counties:

```{r}
#| eval: false
nh_hu <- get_housing_units(
  variables = "housing_units_per_1k_people",
  geoids    = grep("^33", unique(df$geoid), value = TRUE),
  years     = 2010:2024
)
dplyr::glimpse(nh_hu)
```

### Rural vs. Nonrural

```{r}
#| eval: false
library(cori.charts)
library(ggplot2)
library(ruraldefinitions)
library(dplyr)

load_fonts()

rural_avg <- df |>
  filter(variable == "housing_units_per_1k_people") |>
  left_join(select(cbsa_2023, geoid, is_rural), by = "geoid") |>
  filter(!is.na(is_rural), !is.na(value), !is.na(agg_var)) |>
  group_by(year, is_rural) |>
  summarise(
    value = sum(value * agg_var, na.rm = TRUE) / sum(agg_var, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(rural_avg, aes(x = year, y = value, color = is_rural)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_color_manual(
    values = c("Rural" = "#2F6E9B", "Nonrural" = "#7EBDC2"),
    labels = c("Rural" = "Rural counties", "Nonrural" = "Nonrural counties")
  ) +
  scale_x_continuous(breaks = seq(2000, 2024, by = 5)) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
  theme_cori() +
  theme(legend.position = "bottom") +
  labs(
    title    = "Rural counties carry more housing units per capita than nonrural",
    subtitle = "Population-weighted average housing units per 1,000 population, 2000\u20132024",
    x        = NULL, y = NULL, color = NULL,
    caption  = "Source: CORI analysis of U.S. Census Bureau Housing Unit Estimates."
  )
```

### County Spotlight: Grafton County, NH

```{r}
#| eval: false
grafton <- get_housing_units(
  variables = "housing_units_per_1k_people",
  geoids    = c("33009", "33", "00")
) |>
  mutate(
    group = case_when(
      geoid == "33009" ~ "Grafton County, NH",
      geoid == "33"    ~ "New Hampshire",
      geoid == "00"    ~ "United States"
    ),
    group = factor(group, levels = c("Grafton County, NH", "New Hampshire", "United States"))
  )

ggplot(grafton, aes(x = year, y = value,
                    color = group, linetype = group, linewidth = group)) +
  geom_line(na.rm = TRUE) +
  geom_point(size = 2.5, na.rm = TRUE) +
  scale_color_manual(values = c(
    "Grafton County, NH" = "#2F6E9B",
    "New Hampshire"      = "#7EBDC2",
    "United States"      = "#9DA7B0"
  )) +
  scale_linetype_manual(values = c(
    "Grafton County, NH" = "solid",
    "New Hampshire"      = "dashed",
    "United States"      = "dotted"
  )) +
  scale_linewidth_manual(values = c(
    "Grafton County, NH" = 1.2,
    "New Hampshire"      = 0.8,
    "United States"      = 0.8
  )) +
  scale_x_continuous(breaks = seq(2000, 2024, by = 5)) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
  theme_cori() +
  theme(legend.position = "bottom") +
  labs(
    title    = "Housing units per 1,000 residents \u2014 Grafton County, NH",
    subtitle = "Compared to New Hampshire and the United States, 2000\u20132024",
    x        = NULL, y = NULL, color = NULL, linetype = NULL, linewidth = NULL,
    caption  = paste0(
      "Source: CORI analysis of U.S. Census Bureau Housing Unit Estimates.\n",
      "Population denominator from Census Population Estimates Program via cori.data.pep."
    )
  )
```

## Related

-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `cori.data.hu`
-   **Package:** [cori.data.bps](/packages/cori-data-bps/) — building
    permits measure new supply added annually; pairs with HU for stock
    vs. flow analysis
-   **Package:** [cori.data.pep](/packages/cori-data-pep/) — population
    denominators used to compute housing units per capita

::: callout-note
**Data attribution:** Housing Unit Estimates are produced by the U.S.
Census Bureau. When citing analysis using this package, reference: U.S.
Census Bureau, Housing Unit Estimates,
[www.census.gov/programs-surveys/popest](https://www.census.gov/programs-surveys/popest).
:::
