---
title: "cori.data.hu"
description: "Access and analyze U.S. Census Bureau Housing Unit Estimates at the national, state, and county level from 2020 to present"
author: "Drew Rosebush"
date: "2026-08-26"
categories: ["R Package", "Housing"]
tags: ["Housing Units", "Census", "Housing Stock", "County-Level", "S3", "Tidy Data"]
package:
  name: "cori.data.hu"
  githubUrl: "https://github.com/ruralinnovation/cori.data.hu"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.hu")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["census-housing-units"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.hu` package provides access to the U.S. Census Bureau's
**National, State, and County Housing Unit Totals** — annual estimates
of the total number of housing units in existence, published under the
Population Estimates Program. Data cover 2000 to present and measure
housing **stock** (occupied and vacant units combined), built from the
decennial census baseline plus annual permitted construction,
non-permitted construction, mobile home shipments, and estimated
housing loss — as distinct from construction activity or permitting
flow.

::: callout-note
## What's in the package

-   **Total housing stock** — Measures all existing units, not just new
    construction; pairs with `cori.data.bps` for net-new activity
-   **S3-backed access** — Pre-processed data queryable via DuckDB
    without downloading per-state source files
-   **Three anchored decade series** — 2000s, 2010s, 2020s, each reset
    to its own decennial census baseline
-   **Methodology-aware** — README documents the components-of-change
    model and known limitations for longitudinal use
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.hu")
```

## Quick start

```{r}
#| eval: false

library(cori.data.hu)

# Tidy long-format access
df <- get_housing_units(geography = "county")

# Variable documentation
get_hu_codebook()
```

## Core functions

`get_housing_units()` is the tidy-format entry point:

```{r}
#| eval: false

df <- get_housing_units(geography = "county")

# Housing units per 1,000 population for New Hampshire counties
nh_hu <- get_housing_units(
  variables = "housing_units_per_1k_people",
  geoids    = grep("^33", unique(df$geoid), value = TRUE),
  years     = 2010:2024
)
```

`read_hu_from_s3()` reads pre-processed data directly from S3 via
DuckDB:

```{r}
#| eval: false

read_hu_from_s3()
```

### Components-of-change methodology

The Census Bureau builds housing unit estimates using:

```
2020 Census Housing Units
  + New Residential Construction  (permitted × completion rate + non-permitted)
  + New Mobile Homes              (Manufactured Homes Survey)
  - Housing Units Lost            (AHS-based loss rates + FEMA disaster losses)
= July 1 Annual Housing Unit Estimate
```

A "Vintage" designation covers the estimate window, e.g. Vintage 2024
covers April 1, 2020 – July 1, 2024.

## Common use cases

### Rural vs. nonrural comparison

Housing stock per capita diverges between rural and nonrural counties:

```{r}
#| eval: false

library(dplyr)
library(ruraldefinitions)

df <- get_housing_units(
  geography = "county",
  variables = "housing_units_per_1k_people"
)

rural_avg <- df |>
  left_join(select(cbsa_2023, geoid, is_rural), by = "geoid") |>
  filter(!is.na(is_rural), !is.na(value), !is.na(agg_var)) |>
  group_by(year, is_rural) |>
  summarise(
    value = sum(value * agg_var, na.rm = TRUE) / sum(agg_var, na.rm = TRUE),
    .groups = "drop"
  )
```

CORI's own analysis using this pattern found rural counties carry more
housing units per capita than nonrural counties across the full
2000–2024 series.

### County spotlight

```{r}
#| eval: false

grafton <- get_housing_units(
  variables = "housing_units_per_1k_people",
  geoids    = c("33009", "33", "00")  # county, state, national
)
```

## Data coverage

| Dimension | Coverage |
|---|---|
| **Geography** | National, state, county (5-digit FIPS) |
| **Time period** | 2000–present, as three independent decade series (2000s, 2010s, 2020s), each anchored to its own decennial census baseline |
| **Vintage** | Annual (typically released each May) |
| **Update frequency** | Annual |
| **Source** | Census Bureau Population Estimates Program |

## Known limitations

::: callout-important
## Cross-decade comparisons

Do **not** compute net change across the 2009→2010 or 2019→2020
boundaries. Each decade series is anchored to its own decennial census
baseline — the 2010 and 2020 values are fresh Census counts, not
continuations of the prior series' estimate. Subtracting across these
resets produces a methodological artifact, not a real change in
housing stock. For year-over-year change analysis, restrict to a
single decade series or use `cori.data.bps` permit data as a proxy for
net new construction.
:::

Additional limitations:

-   **No structure type breakdown** in the public release — pair with
    BPS data for single-family vs. multifamily distinctions
-   **No vacancy or occupancy status** — pair with ACS for vacancy rates
-   **Loss rates are static** — derived from 2009/2011 AHS, unchanged
    across all vintages; demolition patterns may be underestimated in
    active redevelopment markets
-   **Non-permitted construction not distributed to the West** —
    methodology gap for Western counties without permit offices
-   **County is the smallest public geography** — subcounty (city/town)
    estimates exist internally at Census but are not released
-   **Estimates, not counts** — all values are modeled; the decennial
    census is the only direct count

## Related resources

### Datasets

-   Census Housing Unit Estimates *(slug: `census-housing-units` — node
    pending)* — Source dataset documentation

### Related packages

-   [cori.data.bps](/packages/cori-data-bps/) — Building permits, a
    proxy for net-new construction within a decade series
-   [cori.data.pep](/packages/cori-data-pep/) — Population estimates
    for per-capita denominators

### External links

-   [Census Housing Unit Estimates Dataset Page](https://www.census.gov/data/datasets/time-series/demo/popest/2020s-total-housing-units.html)
-   [Vintage 2024 Methodology (PDF)](https://www2.census.gov/programs-surveys/popest/technical-documentation/methodology/2020-2024/2024-hu-meth.pdf)
-   [Vintage 2020 Methodology (PDF)](https://www2.census.gov/programs-surveys/popest/technical-documentation/methodology/2010-2020/2020-hu-meth_final.pdf)
-   [Building Permits Survey](https://www.census.gov/construction/bps/)
-   [Survey of Construction](https://www.census.gov/construction/nrc/about_the_surveys/soc.html)
-   [Manufactured Homes Survey](https://www.census.gov/programs-surveys/mhs.html)
-   [American Housing Survey](https://www.census.gov/programs-surveys/ahs/)

::: callout-note
## Dangling references

The following slugs are referenced by this package but do not yet have
nodes in Dataverse. They are intentionally preserved as future content
needs:

-   `dataset/census-housing-units`
:::

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.hu/issues).
