---
title: "cori.data.bps"
description: "Access and analyze U.S. Census Bureau Building Permits Survey (BPS) data covering new residential units authorized from 2000 to present"
author: "Drew Rosebush"
date: "2026-08-26"
categories: ["R Package", "Housing"]
tags: ["BPS", "Building Permits", "Housing", "County-Level", "S3", "Tidy Data"]
package:
  name: "cori.data.bps"
  githubUrl: "https://github.com/ruralinnovation/cori.data.bps"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.bps")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["census-building-permits"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.bps` package provides access to U.S. Census Bureau
Building Permits Survey (BPS) data covering new residential units
authorized from 2000 to present, at the county, state, and national
level.

Data are returned in long format: one row per `geoid / year / variable`.

::: callout-note
## What's in the package

-   **Single entry point** — one function, `get_building_permits()`,
    covers geography, year, and variable filtering
-   **Population-weighted rates** — an `agg_var` column carries the
    population denominator so per-capita rates can be correctly
    weighted when aggregating
-   **S3-backed access** — pre-processed data queried via DuckDB
    without downloading full Census source files
-   **Long panel** — annual data from 2000 to present
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.bps")
```

**Census API key** (required):
```{r}
#| eval: false

# Get one at: http://api.census.gov/data/key_signup.html
Sys.setenv(CENSUS_API_KEY = "your_api_key_here")
```

**S3 access** (for reading processed data):
```{r}
#| eval: false

# cori.data.bps depends on cori.data.s3 for S3 reads.
# Falls back to CORI's credential-vending endpoint if no
# local AWS credentials are configured.
remotes::install_github("ruralinnovation/cori.data.s3")
```

## Quick start

```{r}
#| eval: false

library(cori.data.bps)

# See available variables
get_bps_codebook()

# County-level building permits
df <- get_building_permits(geography = "county")

# State-level data for specific years
state_df <- get_building_permits(geography = "state", years = 2010:2024)
```

## Core function

`get_building_permits()` is the single entry point for this package:

```{r}
#| eval: false

# Returns: geoid | year | variable | value | agg_var
get_building_permits(
  geography = "county",
  years     = NULL,
  geoids    = NULL,
  variables = NULL,
  vintage   = "latest"
)
```

| Parameter | Description |
|---|---|
| `geography` | One of `"county"`, `"state"`, or `"nation"` |
| `years` | Integer vector of years to return (default: all available) |
| `geoids` | FIPS codes to return (overrides `geography` when supplied) |
| `variables` | Variables to return (default: all) |
| `vintage` | Data vintage, e.g. `"2024"` (default: `"latest"`) |

## Key variables

| Variable | Description | Unit |
|---|---|---|
| `building_permits` | New residential units authorized | units |
| `units_per_1k_people` | Units authorized per 1,000 population | units per 1,000 persons |

The `agg_var` column carries `population / 1,000` — the correct
denominator for population-weighted averaging of the rate variable.
"Total" includes single-family, 2-family, 3-to-4-family, and
5-plus-family units.

Use `get_bps_codebook()` for complete variable documentation.

## Filtering and selection

### Filter by geography

```{r}
#| eval: false

# Specific counties by FIPS
nh_counties <- get_building_permits(
  geoids = c("33009", "33011", "33013"),
  years  = 2010:2024
)

# Compare county, state, and national
grafton <- get_building_permits(
  geoids = c("33009", "33", "00")
)
```

### Select specific variables

```{r}
#| eval: false

rates <- get_building_permits(
  geography = "county",
  variables = "units_per_1k_people",
  years     = 2015:2024
)
```

## Common use cases

### Population-weighted averages

The `agg_var` column enables correct population-weighted averaging
rather than a naive mean across counties:

```{r}
#| eval: false

library(dplyr)

df <- get_building_permits(geography = "county")

df |>
  filter(variable == "units_per_1k_people") |>
  group_by(year) |>
  summarise(
    weighted_avg = sum(value * agg_var, na.rm = TRUE) / sum(agg_var, na.rm = TRUE)
  )
```

### Rural vs. nonrural comparison

```{r}
#| eval: false

library(dplyr)
library(ruraldefinitions)

rural_avg <- df |>
  filter(variable == "units_per_1k_people") |>
  left_join(select(cbsa_2023, geoid, is_rural), by = "geoid") |>
  group_by(year, is_rural) |>
  summarise(
    value = sum(value * agg_var, na.rm = TRUE) / sum(agg_var, na.rm = TRUE),
    .groups = "drop"
  )
```

CORI's own analysis using this pattern found nonrural counties
consistently authorize more housing per capita than rural counties
across the full 2000–2024 series.

### County spotlight

The same `geoids` filter works across geography levels in a single
call, useful for comparing a county against its state and the nation:

```{r}
#| eval: false

grafton <- get_building_permits(
  variables = "units_per_1k_people",
  geoids    = c("33009", "33", "00")  # county, state, national
)
```

## Data coverage

| Dimension | Coverage |
|---|---|
| **Source** | U.S. Census Bureau, Building Permits Survey |
| **Time period** | 2000–present |
| **Update frequency** | Annual (release typically May) |
| **Geography** | National (`"00"`), state (2-digit FIPS), county (5-digit FIPS) |
| **Structure types** | Single-family, 2-family, 3-to-4-family, 5-plus-family (total combines all) |

## Related resources

### Datasets

-   [Census Building Permits Survey](/datasets/census-building-permits/) —
    Source dataset documentation

### Related packages

-   [cori.data.pep](/packages/cori-data-pep/) — Population estimates,
    the source of the `agg_var` denominator used for per-capita rates

### External links

-   [Census Building Permits Survey](https://www.census.gov/construction/bps/)
-   [BPS Methodology](https://www.census.gov/construction/bps/methodology.html)

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data.bps)
-   [Full README](https://github.com/ruralinnovation/cori.data.bps/blob/main/README.md)
-   [Introduction Vignette](https://github.com/ruralinnovation/cori.data.bps/blob/main/vignettes/introduction.Rmd)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.bps/issues).
