---
title: cori.data.bps
description: Access and analyze U.S. Census Bureau Building Permits Survey (BPS)
  data covering new residential units authorized from 2000 to present. Tracks
  housing supply growth at the county, state, and national level as an indicator
  of local market conditions and development capacity.
author: Drew Rosebush
date: '2026-08-31'
categories:
  - R Package
  - Housing
tags:
  - BPS
  - Building Permits
  - Housing
  - County-Level
  - S3
  - Tidy Data
  - Rural
package:
  name: cori.data.bps
  githubUrl: https://github.com/ruralinnovation/cori.data.bps
  installCommand: remotes::install_github("ruralinnovation/cori.data.bps")
  status: development
  version: 0.1.0
  maintainer: Drew Rosebush
  featured: false
usesDatasets:
  - census-building-permits
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.bps` package provides access to U.S. Census Bureau
Building Permits Survey (BPS) data at the county, state, and national
level from 2000 to the present. BPS tracks new residential units
authorized by building permits across all structure types — single-unit,
two-to-four unit, and five-or-more unit — providing an annual measure of
how housing supply is growing in response to local conditions.

Building permits are the earliest observable signal of housing supply
responding to conditions on the ground — and in communities where
housing is a binding constraint on growth, that signal matters. Permit
counts are normalized by population using `cori.data.pep` and returned
in the standard `cori.data` long format: one row per `geoid / year /
variable`.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/cori.data.bps")
```

```{r}
#| eval: false
library(cori.data.bps)

# All available variables
get_bps_codebook()

# All counties, all years
get_building_permits(geography = "county")

# New Hampshire counties, permits per 1k population, 2010 onward
get_building_permits(
  variables = "units_per_1k_people",
  geoids    = c("33009", "33011"),
  years     = 2010:2024
)
```

## What This Data Tells Us About Rural

Housing supply is a foundational asset for rural economic development —
and in many communities, a binding constraint on growth strategies that
are otherwise well-designed. CORI's place-based community research has
found communities with genuine competitive assets — skilled institutions,
entrepreneurial infrastructure, strong broadband — where housing
shortfall functions as a ceiling on everything else. Workers cannot
relocate to communities where housing is unavailable or unaffordable,
and population retention becomes harder when housing stock does not keep
pace with demand.

Building permit data captures the rate at which housing supply is
responding to conditions on the ground. An increase in permit activity
indicates that supply is growing in response to demand, and that local
regulatory and market conditions allow it. Where permit activity lags
population or employment growth, the data surfaces a constraint that
deserves attention alongside other economic development investments.

## Introduction

*Adapted from the [Introduction
vignette](https://github.com/ruralinnovation/cori.data.bps/blob/main/vignettes/introduction.Rmd)
on GitHub.*

`cori.data.bps` provides annual building permit data from the U.S.
Census Bureau's Building Permits Survey. Data cover 2000–present and
measure new residential units authorized by building permits across all
structure types.

**Source:** U.S. Census Bureau, Building Permits Survey
**Coverage:** 2000–present, updated annually (release typically May)
**Geography:** County (5-digit FIPS), state (2-digit FIPS), national
(`"00"`)

### Variables

```{r}
#| eval: false
get_bps_codebook() |>
  dplyr::select(variable, label, unit, notes) |>
  knitr::kable()
```

### Reading data

All data are returned in long format: one row per `geoid / year /
variable`.

```{r}
#| eval: false
df <- get_building_permits(geography = "county")
dplyr::glimpse(df)
```

Filter to specific variables, years, or geographies:

```{r}
#| eval: false
nh_bps <- get_building_permits(
  variables = "units_per_1k_people",
  geoids    = grep("^33", unique(df$geoid), value = TRUE),
  years     = 2010:2024
)
dplyr::glimpse(nh_bps)
```

### Rural vs. Nonrural

Housing permit activity diverges between rural and nonrural counties.
The chart below shows population-weighted average units permitted per
1,000 people by rural status using the CBSA 2023 rural definition.

```{r}
#| eval: false
library(cori.charts)
library(ggplot2)
library(ruraldefinitions)
library(dplyr)

load_fonts()

rural_avg <- df |>
  filter(variable == "units_per_1k_people") |>
  left_join(select(cbsa_2023, geoid, is_rural), by = "geoid") |>
  filter(!is.na(is_rural), !is.na(value), !is.na(agg_var)) |>
  group_by(year, is_rural) |>
  summarise(
    value = sum(value * agg_var, na.rm = TRUE) / sum(agg_var, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(rural_avg, aes(x = year, y = value, color = is_rural)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_color_manual(
    values = c("Rural" = "#2F6E9B", "Nonrural" = "#7EBDC2"),
    labels = c("Rural" = "Rural counties", "Nonrural" = "Nonrural counties")
  ) +
  scale_x_continuous(breaks = seq(2000, 2024, by = 5)) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
  theme_cori() +
  theme(legend.position = "bottom") +
  labs(
    title    = "Nonrural counties consistently authorize more housing per capita",
    subtitle = "New residential units authorized per 1,000 population, 2000\u20132024",
    x        = NULL, y = NULL, color = NULL,
    caption  = "Source: CORI analysis of U.S. Census Bureau Building Permits Survey."
  )
```

### County Spotlight: Grafton County, NH

```{r}
#| eval: false
grafton <- get_building_permits(
  variables = "units_per_1k_people",
  geoids    = c("33009", "33", "00")
) |>
  mutate(
    group = case_when(
      geoid == "33009" ~ "Grafton County, NH",
      geoid == "33"    ~ "New Hampshire",
      geoid == "00"    ~ "United States"
    ),
    group = factor(group, levels = c("Grafton County, NH", "New Hampshire", "United States"))
  )

ggplot(grafton, aes(x = year, y = value,
                    color = group, linetype = group, linewidth = group)) +
  geom_line(na.rm = TRUE) +
  geom_point(size = 2.5, na.rm = TRUE) +
  scale_color_manual(values = c(
    "Grafton County, NH" = "#2F6E9B",
    "New Hampshire"      = "#7EBDC2",
    "United States"      = "#9DA7B0"
  )) +
  scale_linetype_manual(values = c(
    "Grafton County, NH" = "solid",
    "New Hampshire"      = "dashed",
    "United States"      = "dotted"
  )) +
  scale_linewidth_manual(values = c(
    "Grafton County, NH" = 1.2,
    "New Hampshire"      = 0.8,
    "United States"      = 0.8
  )) +
  scale_x_continuous(breaks = seq(2000, 2024, by = 5)) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) +
  theme_cori() +
  theme(legend.position = "bottom") +
  labs(
    title    = "Grafton County, NH\nNew residential units authorized per 1,000 residents",
    subtitle = "Compared to New Hampshire and the United States, 2000\u20132024",
    x        = NULL, y = NULL, color = NULL, linetype = NULL, linewidth = NULL,
    caption  = paste0(
      "Source: CORI analysis of U.S. Census Bureau Building Permits Survey.\n",
      "Population denominator from Census Population Estimates Program via cori.data.pep."
    )
  )
```

## Related

-   **Dataset:** [Census Building Permits](/datasets/census-building-permits/)
-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `cori.data.bps`
-   **Package:** [cori.data.pep](/packages/cori-data-pep/) — population
    denominators used to compute permits per 1,000 people
-   **Package:** [cori.data.hu](/packages/cori-data-hu/) — total housing
    stock; pairs with BPS to distinguish new supply from existing stock
-   **Project:** [Rural Economic
    Outlook](/projects/rural-economic-outlook/) — BPS is an input to
    CORI's housing supply analysis

::: callout-note
**Data attribution:** Building Permits Survey data are produced by the
U.S. Census Bureau. When citing analysis using this package, reference:
U.S. Census Bureau, Building Permits Survey,
[www.census.gov/construction/bps](https://www.census.gov/construction/bps).
:::
