---
title: cori.data.fcc
description: Functions to get and process FCC data.
author: Olivier Leroy
date: '2026-08-26'
categories:
  - R Package
  - Broadband
tags:
  - broadband
  - fcc
  - nbm
  - national-broadband-map
  - form-477
  - census-block
  - isp
package:
  name: cori.data.fcc
  githubUrl: https://github.com/ruralinnovation/cori.data.fcc
  installCommand: remotes::install_github("ruralinnovation/cori.data.fcc")
  status: stable
  version: 1.0.0
  maintainer: Olivier Leroy
  featured: false
usesDatasets:
  - fcc-broadband
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The goal of `cori.data.fcc` is to facilitate the discovery, analysis, and use of FCC public data releases: the [National Broadband Map](https://broadbandmap.fcc.gov/home) (NBM) and [Form 477](https://www.fcc.gov/general/broadband-deployment-data-fcc-form-477).

NBM and Form 477 are two eras of the same underlying question — where is broadband actually available. Form 477 required facilities-based providers to file twice-yearly, census-block-level (2010 vintage) coverage reports from December 2014 through December 2021. NBM launched in November 2022 as its successor, moving from census-block reporting to individual location-based reporting derived from CostQuest's "Fabric" locations dataset, with two major releases a year (June and December) plus incremental updates roughly every two weeks to incorporate challenges. The FCC does not publish a changelog between releases.

The package provides S3-backed access to NBM data — both raw parquet files by county and a CORI-opinionated Census-block-level summary — as well as Form 477 broadband deployment data. A built-in data dictionary and ISP provider reference table round out the package.

::: callout-note
## What's in the package

- **S3-backed NBM access via DuckDB** — Raw NBM parquet files queryable by county without local download, powered by DuckDB.
- **CORI-opinionated block layer** — 21-column Census-block summary with technology type breakdowns (fiber, cable, copper, fixed wireless, etc.) and BEAD location counts.
- **FRN-based coverage queries** — Retrieve all Census blocks served by a specific ISP using their FCC Registration Number.
- **Form 477 multi-year data** — State-level broadband deployment records, December 2014 through December 2021.
- **Integrated data dictionary** — Per-dataset codebook covering `f477`, `nbm_block`, and `nbm_raw` field descriptions.
- **ISP provider reference** — `fcc_provider` data frame with 4,456 records linking provider names, FRNs, and numeric provider IDs.
:::

## Installation

::: panel-tabset
### GitHub

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.fcc")
```
:::

## Quick start

```{r}
#| eval: false

library(cori.data.fcc)

# Get CORI-opinionated NBM block-level data for a county
nbm_bl <- get_nbm_bl(geoid_co = "47051")

# Get raw NBM parquet data for a county
guilford_cty <- get_nbm_county_raw(geoid_co = "37081")

# Get all blocks served by a specific ISP (by FRN)
skymesh <- get_frn_nbm_bl("0027136753")

# Get Form 477 data for a state
f477_vt <- get_f477("VT")
```

## Discovering available data

NBM ships a new release every few weeks. Before pulling data for a
specific county, check what releases and files actually exist:

```{r}
#| eval: false

# What releases are available?
release <- get_nbm_release()

# What files are available within a release?
nbm <- get_nbm_available()

# Filter to what you need, e.g. Fixed Broadband / Nationwide / a specific release
nbm |>
  dplyr::filter(
    release       == "June 30, 2023",
    data_type     == "Fixed Broadband",
    data_category == "Nationwide"
  )
```

## Core functions

::: panel-tabset

### Data access

| Function | Description |
|----------|-------------|
| `get_nbm_bl()` | CORI-opinionated NBM Census-block summary by county (21 columns) |
| `get_nbm_county()` | NBM county-level aggregated data |
| `get_nbm_county_raw()` | Raw NBM parquet data for a county and release via S3 |
| `get_frn_nbm_bl()` | All Census blocks where a given FRN reported service |
| `get_f477()` | Form 477 broadband deployment data by state |

### Utility

| Function | Description |
|----------|-------------|
| `get_nbm_available()` | List available NBM releases in the S3 store |
| `get_nbm_release()` | NBM release metadata |
| `dl_nbm()` | Download NBM source files |
| `check_frn()` | Validate an FCC Registration Number (FRN) |
| `get_fcc_dictionary()` | Codebook for one dataset — pass `"nbm_raw"`, `"nbm_block"`, or `"f477"` |

:::

## Data sources

| Dataset | Role |
|---------|------|
| [FCC Broadband](/datasets/fcc-broadband/) | Primary |

## Use cases at CORI

This package is used in:

- Analyzing broadband availability at the Census-block level for a county or region.
- Identifying BEAD-eligible locations using CORI-processed NBM data.
- Querying coverage footprint for a specific ISP by FRN.
- Tracking broadband deployment trends over time using Form 477 filings (2014–2021).
- Building broadband gap analyses using technology-type breakdowns.

See [Democratizing analytics on FCC's (big) data](/blog/12_fcc-data/)
for more on the different ways CORI uses this data.

::: callout-tip
## Optimization tips

Ensure the latest versions of `duckdb` and `waldo` are installed before use. The raw NBM data (parquet via S3) can be large — queries against `get_nbm_county_raw()` on dense counties may return over one million rows.
:::

::: callout-warning
## Known limitations

- The package accesses data stored in CORI S3 storage; AWS credentials or network access may be required.
- NBM data reflects ISP-reported coverage and is subject to the known limitations of self-reported broadband availability data.
- Form 477 collection ended in December 2021, superseded by NBM — it is not being updated further.
- CORI's Form 477 ingestion keeps only the last release of each filing period (assumed most consistent) and only *Fixed without satellite* data — satellite service is not included.
- FRNs should be a zero-padded 10-digit code, but this has not been consistently enforced across historical filings; `check_frn()` validates format but cannot fix bad source data.
- Data is provided "as is" — CORI does not remove rows with dubious values. That judgment is left to the analyst.
:::

## Documentation

- [GitHub Repository](https://github.com/ruralinnovation/cori.data.fcc)
- [Full README](https://github.com/ruralinnovation/cori.data.fcc/blob/main/README.md)
- [Function Documentation](https://github.com/ruralinnovation/cori.data.fcc/tree/main/man)
- [Vignettes](https://github.com/ruralinnovation/cori.data.fcc/tree/main/vignettes)
