---
title: cori.data.fcc
description: Access FCC National Broadband Map (NBM) and Form 477 data at the county
  and census block level. Includes CORI-processed S3-backed datasets and tools for
  analyzing broadband infrastructure availability across rural and nonrural geographies.
author: John Hall
date: '2026-08-31'
categories:
  - R Package
  - Broadband
tags:
  - FCC
  - NBM
  - Broadband
  - National Broadband Map
  - Block-Level
  - County-Level
  - S3
  - Rural
package:
  name: cori.data.fcc
  githubUrl: https://github.com/ruralinnovation/cori.data.fcc
  installCommand: remotes::install_github("ruralinnovation/cori.data.fcc")
  status: stable
  version: 1.0.0
  maintainer: John Hall
  featured: false
usesDatasets:
  - fcc-broadband
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.fcc` package provides access to two FCC broadband
datasets: the **National Broadband Map (NBM)**, launched in November
2022, and the legacy **Form 477** data covering December 2014 through
December 2021. Together they offer a continuous picture of broadband
infrastructure availability across the United States — from the
census-block-level reporting of Form 477 to the location-level precision
of the NBM.

The NBM is the primary dataset for current analysis. It captures what
internet services are available at individual broadband serviceable
locations (BSLs) — residential and business addresses — as reported by
ISPs twice yearly. CORI makes the NBM available at two analytical levels:
raw location-level data by county (`get_nbm_county_raw()`), and a
census-block summary (`get_nbm_bl()`) that aggregates service counts by
technology type and speed threshold. The block-level summary is the
standard entrypoint for county- and community-level broadband analysis.

Form 477 provides historical context — particularly for tracking how
rural broadband coverage has changed over the decade prior to the NBM.
CORI's ingestion focuses on fixed broadband (excluding satellite) and
applies consistent versioning decisions to produce a comparable
longitudinal dataset.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/cori.data.fcc")
```

```{r}
#| eval: false
library(cori.data.fcc)

# Census block-level NBM summary for a county
nbm_bl <- get_nbm_bl(geoid_co = "33009")

# County-level NBM summary
nbm_co <- get_nbm_county("33009")

# Raw location-level NBM data
nbm_raw <- get_nbm_county_raw(geoid_co = "33009")

# Form 477 — state data across all available years
f477_nh <- get_f477("NH")

# Data dictionary
get_fcc_dictionary("nbm_block")
```

## What This Data Tells Us About Rural

Broadband infrastructure is a foundational asset for rural participation
in the modern economy — enabling remote work, business formation,
telehealth, and education in communities that geography once isolated
from these opportunities. The FCC National Broadband Map provides the
current county-level picture of where infrastructure exists and at what
speeds, making it possible to track the closing of the rural broadband
gap over time.

Research shows that high-speed internet improves economic outcomes in
rural communities through increased business activity and broader
economic development measures including productivity, jobs, and income.
The NBM data makes it possible to track where rural broadband investment
is occurring, where gaps persist, and how coverage has evolved over time.

## National Broadband Map (NBM)

*Adapted from the [NBM
vignette](https://github.com/ruralinnovation/cori.data.fcc/blob/main/vignettes/NBM.Rmd)
on GitHub.*

### What is the NBM?

The FCC's National Broadband Map captures broadband availability at the
level of individual broadband serviceable locations (BSLs) — every
residential and business address where fixed internet service is or can
be installed. For each location, the NBM records which providers serve
it, what technologies they use, and their maximum advertised download and
upload speeds.

The NBM publishes two major releases per year (June and December), with
interim updates every two weeks to incorporate challenge data. CORI
ingests each release into S3 and maintains the full history so analyses
can be compared across time.

### What is a Broadband Serviceable Location (BSL)?

A BSL is defined as "a business or residential location in the United
States at which fixed broadband internet access service is, or can be,
installed." The exact coordinates of each location are part of the
underlying Fabric dataset maintained by CostQuest Associates — within
the Broadband Availability data, locations are linked to a 2020 Census
block.

### Accessing NBM data

The package provides three levels of access:

**1. County-level summary** — aggregated counts per county per release:

```{r}
#| eval: false
nbm_co <- get_nbm_county("33009")
dplyr::glimpse(nbm_co)
```

**2. Census block-level** — CORI-opinionated aggregation with technology
type and speed threshold counts (21 columns), one row per block:

```{r}
#| eval: false
nbm_bl <- get_nbm_bl(geoid_co = "33009")
dplyr::glimpse(nbm_bl)
```

**3. Raw location-level** — the full NBM records for all locations in a
county as reported by ISPs:

```{r}
#| eval: false
nbm_raw <- get_nbm_county_raw(geoid_co = "33009")
dplyr::glimpse(nbm_raw)
```

The block-level summary (`get_nbm_bl()`) is the recommended starting
point for county and community analysis — it aggregates to technology
type and speed thresholds most relevant for rural broadband gap work
(25/3, 100/20, 100/100 Mbps).

### Checking available releases

```{r}
#| eval: false
# List all available NBM release dates
get_nbm_release()

# See what files are available
get_nbm_available()
```

### Looking up providers by FRN

The FCC Registration Number (FRN) uniquely identifies each ISP. To find
all census blocks served by a specific provider:

```{r}
#| eval: false
# Blocks served by a specific provider
get_frn_nbm_bl("0027136753")

# Validate an FRN format
check_frn("0027136753")
```

A reference table of provider names, FRNs, and provider IDs is bundled
as `fcc_provider` (4,456 records).

### Data dictionary

```{r}
#| eval: false
get_fcc_dictionary("nbm_block")  # block-level variables
get_fcc_dictionary("nbm_raw")    # raw NBM variables
```

::: callout-note
**Attribution:** Broadband availability data from the FCC Broadband Data
Collection is offered free and not subject to copyright restriction. When
using this data, please attribute as: *Source data: FCC Broadband Data
Collection.* The underlying Fabric location data is owned by CostQuest
Associates, Inc. See the [NBM about
page](https://broadbandmap.fcc.gov/about) for full licensing terms.
:::

## Form 477

*Adapted from the [Form 477
vignette](https://github.com/ruralinnovation/cori.data.fcc/blob/main/vignettes/f477.Rmd)
on GitHub.*

Form 477 is the FCC's legacy broadband reporting program, covering
December 2014 through December 2021. Providers were required to file
twice yearly on where they offered internet access service at speeds
exceeding 200 kbps, at the census block level (2010 vintage). Unlike the
NBM, Form 477 reported at the block level rather than the location level
— a provider needed only to indicate they served at least one location in
a block, not what share of the block was covered.

CORI's ingestion keeps only the final version of each filing period
(assumed to be the most consistent release) and focuses on fixed
broadband only, excluding satellite. The data is provided as-is — values
are not filtered or corrected, leaving those judgments to the analyst.

```{r}
#| eval: false
# Form 477 for New Hampshire — all available years
f477_nh <- get_f477("NH")
dplyr::glimpse(f477_nh)

# Data dictionary
get_fcc_dictionary("f477")
```

Form 477 is most useful for historical analysis — understanding broadband
coverage before the NBM era, or examining pre/post-NBM coverage trends
in rural counties over a decade of change.

::: callout-note
**Attribution:** Broadband data from FCC Form 477 is offered free and
not subject to copyright restriction. Please attribute as: *Source data:
FCC Form 477.*
:::

## Related

-   **Dataset:** [FCC Broadband](/datasets/fcc-broadband/) — source
    dataset documentation
-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `cori.data.fcc`
-   **Project:** [Rural Economic
    Outlook](/projects/rural-economic-outlook/) — broadband coverage is
    a core input to CORI's multi-dataset economic analysis
-   **Post:** [Democratizing analytics on FCC's (big)
    data](/posts/12_fcc-data/) — more on how CORI uses this data
