---
title: "cori.data.rin"
description: "County-level service area crosswalk for Rural Innovation Network (RIN) communities, updated via a targets pipeline from Comms-maintained Google Sheets"
author: "John Hall"
date: "2026-05-12"
categories: ["R Package", "Network Data", "Rural Development"]
tags: ["RIN", "Rural Innovation Network", "Community Data", "Service Areas", "County-Level", "S3"]
package:
  name: "cori.data.rin"
  githubUrl: "https://github.com/ruralinnovation/cori.data.rin"
  installCommand: 'devtools::install_github("ruralinnovation/cori.data.rin")'
  status: "stable"
  version: "1.0.1"
  maintainer: "John Hall"
  featured: false
usesDatasets: ["rin-communities"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.rin` package exports `rin_service_areas` — the canonical
mapping from Rural Innovation Network (RIN) community names to their
associated U.S. counties, keyed by 5-digit FIPS code. This is the
authoritative county crosswalk used across CORI tools whenever RIN
geography needs to be joined to county-level data.

::: callout-note
## Why Use This Package

-   **Canonical RIN geography** — single source of truth for which
    counties belong to each RIN community
-   **FIPS-ready** — `geoid_co` is a standard 5-digit county FIPS code
    for direct joins to ACS, BDS, COG, or any other county-level dataset
-   **Primary county flag** — distinguishes a community's primary county
    from secondary service area counties
-   **Always current** — MDA refreshes the dataset from the
    Comms-maintained Google Sheet whenever community boundaries change
:::

## Installation

```{r}
#| eval: false

devtools::install_github("ruralinnovation/cori.data.rin")
```

## Quick Start

```{r}
#| eval: false

library(cori.data.rin)
library(dplyr)

# Load the county-level service area crosswalk
rin_service_areas

# How many counties does each RIN community serve?
rin_service_areas |>
  count(rin_community, name = "n_counties") |>
  arrange(desc(n_counties))

# Primary counties only
rin_service_areas |>
  filter(primary_county_flag == "Yes")

# Join RIN community membership onto another county dataset
county_data |>
  left_join(rin_service_areas, by = c("geoid" = "geoid_co"))
```

## Data Structure

### `rin_service_areas`

A data frame with one row per community–county pair:

| Column | Type | Description |
|---|---|---|
| `geoid_co` | character | County FIPS (5-digit: STATEFP + COUNTYFP) |
| `rin_community` | character | CORI name for the RIN community |
| `county` | character | County name |
| `primary_county_flag` | character | `"Yes"` / `"No"` — primary county for this community |
| `data_run_date` | Date | Date the dataset was last refreshed from the source sheet |

## Data Coverage

| Dimension | Coverage |
|---|---|
| **Geography** | County (5-digit FIPS) |
| **Unit** | One row per RIN community × county pair |
| **Source** | Comms-maintained Google Sheet, updated as communities join or expand |
| **Update Frequency** | On demand — MDA runs `targets::tar_make()` after Comms updates the sheet |

## Use Cases

-   **County-level joins** — Attach ACS, BDS, COG, or any other
    county-level dataset to RIN communities via `geoid_co`
-   **Network coverage analysis** — Identify which counties are served
    by one or more RIN communities, and which communities share counties
-   **Reporting and mapping** — Provide the spatial layer for RIN
    community dashboards and the [RIN map frontend](https://github.com/ruralinnovation/rin-map)
-   **Agent knowledge base** — Supplies the geographic context layer for
    CORI's AI agent to answer questions about RIN community locations

## Data Refresh (Admin)

The dataset is generated by a `targets` pipeline maintained by MDA.
After Comms updates the [source Google Sheet](https://docs.google.com/spreadsheets/d/1fDGclR7GqncaLdkEipOJLLwDn3h4gYHVRcJtP_K1wcU):

```{r}
#| eval: false

# From the cori.data.rin project root
targets::tar_make()
```

This rebuilds `rin_service_areas`, writes `data/rin_map.json`, and
uploads outputs to S3 (`cori-risi-apps` for the RIN map frontend,
`cori.agent.kb` for the agent knowledge base). Once verified on dev,
the updated files are copied to the production S3 prefix.

## Related Resources

### Datasets

-   RIN Communities *(slug: `rin-communities` — node pending)* —
    Community and service area data maintained by Comms in Monday.com
    and Google Sheets

### External Links

-   [RIN Map Repository](https://github.com/ruralinnovation/rin-map)
-   [Monday.com Board](https://ruralinnovation-group.monday.com/boards/6951894369)

::: callout-note
## Dangling references

The following slugs are referenced by this package but do not yet have
nodes in Dataverse. They are intentionally preserved as future content
needs:

-   `dataset/rin-communities`
:::

---

**Contact:** John Hall (john.hall@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.rin/issues).
