---
title: "cori.data.s3"
description: "CORI functions used by cori.data.* packages to connect to S3"
author: "John Hall"
date: "2026-08-26"
categories: ["R Package", "Data Infrastructure"]
tags: ["S3", "AWS", "DuckDB", "Credentials", "Internal"]
package:
  name: "cori.data.s3"
  githubUrl: "https://github.com/ruralinnovation/cori.data.s3"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.s3")'
  status: "stable"
  version: "1.0.0"
  maintainer: "John Hall"
  featured: false
usesDatasets: []
usesResources: ["monitoring-s3-access"]
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

`cori.data.s3` is the shared S3 access layer that every `cori.data.*`
package depends on. It is not a data package itself — it provides the
DuckDB connection, credential resolution, and object read/write
functions the other packages call underneath. These functions
previously lived in the `cori.db` package and were split out here.

::: callout-note
## What's in the package

-   **Local-then-vended credentials** — uses local AWS credentials
    when configured (env vars or `~/.aws/credentials`); otherwise
    falls back to temporary, read-only credentials from CORI's
    credential-vending endpoint, no local AWS setup required
-   **DuckDB S3 connections** — `connect_to_s3()` opens a DuckDB
    connection pre-configured for S3, so parquet files can be queried
    directly with SQL
-   **Read and write object functions** — get, list, read, and write
    individual objects or recursive directories
-   **Writes require local credentials** — vended credentials are
    read-only, so `put_s3_object()`, `put_s3_objects_recursive()`,
    `write_s3_object()`, and `list_s3_buckets()` always need local AWS
    setup
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.s3")
```

## Quick start

```{r}
#| eval: false

library(cori.data.s3)

con <- connect_to_s3("cori.data.bds")

DBI::dbGetQuery(con,
  "SELECT * FROM read_parquet('s3://cori.data.bds/**/*.parquet')")

DBI::dbDisconnect(con, shutdown = TRUE)
```

`has_local_aws_credentials()` checks whether local credentials are
configured, without attempting a connection.

## Core functions

| Function | Description |
|---|---|
| `connect_to_s3()` | Open a DuckDB connection configured for S3 access |
| `has_local_aws_credentials()` | Check whether local AWS credentials are configured |
| `get_s3_object()` | Download a single S3 object |
| `list_s3_objects()` | List objects under a bucket/prefix |
| `list_s3_buckets()` | List accessible buckets (requires local credentials) |
| `read_s3_object()` | Read an S3 object directly into memory |
| `write_s3_object()` | Write an object to S3 (requires local credentials) |
| `put_s3_object()` | Upload a single local file to S3 (requires local credentials) |
| `put_s3_objects_recursive()` | Upload a local directory tree to S3 (requires local credentials) |
| `sync_s3_to_local()` | Sync an S3 prefix to a local directory |
| `set_aws_credentials()` | Set AWS credentials for the current session |

## Credential resolution

Reads (`get_s3_object()`, `list_s3_objects()`, `read_s3_object()`) try
local AWS credentials first, then fall back to short-lived, read-only
credentials from CORI's credential-vending endpoint — no local AWS
setup required for read-only work. Writes and `list_s3_buckets()`
always require local credentials, since vended credentials cannot
write.

For monitoring who is actually using vended vs. local credentials
against a bucket, see the related resource below — it parses S3 server
access logs into per-caller download charts.

## Related resources

-   [Monitoring S3 access](/resources/monitoring-s3-access/) — Turns
    S3 server access logs into charts of who is downloading what, by
    bucket and by credential type

### Related packages

Every `cori.data.*` data package depends on `cori.data.s3` for its S3
access:

-   [cori.data.bds](/packages/cori-data-bds/)
-   [cori.data.bfs](/packages/cori-data-bfs/)
-   [cori.data.bps](/packages/cori-data-bps/)
-   [cori.data.fcc](/packages/cori-data-fcc/)
-   [cori.data.hu](/packages/cori-data-hu/)
-   [cori.data.pep](/packages/cori-data-pep/)
-   [cori.data.qcew](/packages/cori-data-qcew/)

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data.s3)
-   [Full README](https://github.com/ruralinnovation/cori.data.s3/blob/main/README.md)

---

**Contact:** John Hall (john.hall@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.s3/issues).
