---
title: cori.data.pep
description: Access and analyze U.S. Census Bureau Population Estimates Program
  (PEP) data at the county level from 2000 to present, including population totals
  and components of change (births, deaths, domestic and international migration).
author: Drew Rosebush
date: '2026-08-31'
categories:
  - R Package
  - Demographics
tags:
  - PEP
  - Population
  - Demographics
  - Components of Change
  - Migration
  - County-Level
  - S3
  - Tidy Data
  - Rural
package:
  name: cori.data.pep
  githubUrl: https://github.com/ruralinnovation/cori.data.pep
  installCommand: remotes::install_github("ruralinnovation/cori.data.pep")
  status: development
  version: 0.1.0
  maintainer: Drew Rosebush
  featured: false
usesDatasets:
  - census-population-estimates
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.pep` package provides access to U.S. Census Bureau
Population Estimates Program (PEP) data at the county, state, and
national level from 2000 to the present. The package covers both
population totals and components of change — births, deaths, domestic
migration, and international migration — enabling researchers to
understand not just whether a community is growing or declining, but
why.

Disaggregating population change into its components — rather than
reading net headcounts alone — is essential for understanding what is
actually happening in a rural community: whether it is losing working-age
residents while gaining retirees, or building through in-migration while
natural change is negative. PEP is also the population denominator used
across the `cori.data` ecosystem — `cori.data.bps` and `cori.data.bfs`
both normalize their counts against PEP population estimates. All data
are returned in the standard `cori.data` long format: one row per
`geoid / year / variable`.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/cori.data.pep")
```

```{r}
#| eval: false
library(cori.data.pep)

# All available variables
get_pep_codebook()

# Total population — all counties, all years
get_population(geography = "county", variables = "population")

# Components of change for West Virginia counties
get_population_change(
  variables = c("domestic_migration", "international_migration", "net_migration"),
  geoids    = grep("^54", unique(get_population(geography = "county")$geoid), value = TRUE),
  years     = 2010:2025
)
```

## What This Data Tells Us About Rural

Population change reflects how people are responding to conditions on
the ground — economic opportunity, quality of life, housing access,
community character. It is often treated as a simple headline indicator,
but the headline number obscures as much as it reveals. A community
losing population through out-migration while also seeing natural
increase is in a different situation than one losing population through
both channels simultaneously. A community attracting working-age
in-migrants is building differently than one growing through retiree
in-migration or international immigration, each of which carries
different implications for the labor market, consumer base, and service
demands.

The Population Estimates Program data supports exactly this kind of
disaggregated reading. Components of change — births, deaths, domestic
migration, international migration — allow researchers and community
leaders to understand *why* population is moving, not just whether it
is. CORI's community analysis found that net population numbers
frequently masked offsetting flows: communities experiencing apparent
stability were in some cases losing working-age residents while gaining
older adults — a pattern with significant implications for workforce,
entrepreneurship, and long-run fiscal capacity.

The in-migration of older adults is an underrecognized asset for rural
communities. Research suggests that individuals migrating to rural areas
later in life bring human, financial, and business capital — their
experience and resources can stimulate local economies, support
entrepreneurship, and deepen community networks. Population dynamics,
read through the full components of change rather than the net headline,
reveal which assets a community is gaining and which it is at risk of
losing.

## Introduction

*Adapted from the [Introduction
vignette](https://github.com/ruralinnovation/cori.data.pep/blob/main/vignettes/introduction.Rmd)
on GitHub.*

`cori.data.pep` provides county-level population estimates from the U.S.
Census Bureau's Population Estimates Program. Data cover 2000–present at
annual frequency.

### Variables

```{r}
#| eval: false
get_pep_codebook() |>
  dplyr::select(variable, label, coverage = notes) |>
  knitr::kable()
```

### Reading data

All data are returned in long format: one row per `geoid / year /
variable`.

```{r}
#| eval: false
# Total population, all counties, all years
pop <- get_population(geography = "county", variables = "population")
dplyr::glimpse(pop)
```

Filter to specific variables, years, or counties:

```{r}
#| eval: false
wv_mig <- get_population_change(
  variables = c("domestic_migration", "international_migration", "net_migration"),
  geoids    = grep("^54", unique(pop$geoid), value = TRUE),
  years     = 2010:2025
)
dplyr::glimpse(wv_mig)
```

### County Spotlight: Grafton County, NH

```{r}
#| eval: false
library(ggplot2)
library(cori.charts)

load_fonts()

grafton_pop <- get_population(
  variables = "population",
  geoids    = "33009",
  years     = 2000:2025
)

ggplot(grafton_pop, aes(x = year, y = value)) +
  geom_line(color = cori_colors[["Emerald"]], linewidth = 1.2) +
  geom_point(color = cori_colors[["Emerald"]], size = 2) +
  scale_x_continuous(breaks = seq(2000, 2025, by = 5)) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme_cori() +
  labs(
    title    = "Grafton County, NH population, 2000\u20132025",
    subtitle = "Total resident population estimate",
    x        = NULL,
    y        = NULL,
    caption  = "Source: CORI analysis of U.S. Census Bureau Population Estimates Program (PEP)."
  )
```

## Related

-   **Dataset:** [Census Population
    Estimates](/datasets/census-population-estimates/)
-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `cori.data.pep`
-   **Package:** [cori.data.qcew](/packages/cori-data-qcew/) — employment
    data; pairs with PEP for employment rate analysis
-   **Package:** [cori.data.bps](/packages/cori-data-bps/) — building
    permits use PEP population as denominator
-   **Package:** [cori.data.bfs](/packages/cori-data-bfs/) — business
    applications use PEP population as denominator
-   **Project:** [Rural Economic
    Outlook](/projects/rural-economic-outlook/) — PEP is a core input to
    CORI's demographic analysis

::: callout-note
**Data attribution:** Population Estimates Program data are produced by
the U.S. Census Bureau. When citing analysis using this package,
reference: U.S. Census Bureau, Population Estimates Program,
[www.census.gov/programs-surveys/popest](https://www.census.gov/programs-surveys/popest).
:::
