---
title: "cori.data.pep"
description: "Access and analyze U.S. Census Bureau Population Estimates Program (PEP) data at the county level from 2000 to present, including population totals and components of change"
author: "Drew Rosebush"
date: "2026-08-26"
categories: ["R Package", "Demographics"]
tags: ["PEP", "Population", "Migration", "County-Level", "S3", "Tidy Data", "Rural"]
package:
  name: "cori.data.pep"
  githubUrl: "https://github.com/ruralinnovation/cori.data.pep"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.pep")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["census-population-estimates", "census-pep-components"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.pep` package provides access to U.S. Census Bureau
Population Estimates Program (PEP) data at the county level from 2000
to present. It covers eight variables including total population,
working-age population (16+), and a full decomposition of population
change into natural change (births minus deaths) and net migration
(domestic and international). Data span three Census vintages,
stitched into a single continuous series.

::: callout-note
## What's in the package

-   **Eight variables** — Total population, pop 16+, births, deaths,
    natural change, domestic migration, international migration, net
    migration
-   **Continuous panel from 2000** — Spans three decennial vintages
    (2000–2009, 2010–2019, 2020–present), stitched into a single
    consistent series
-   **S3-backed access** — Pre-processed data available via DuckDB for
    fast county-level queries
-   **Standard long format** — One row per `geoid / year / variable`,
    ready for tidy analysis
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.pep")
```

## Quick start

```{r}
#| eval: false

library(cori.data.pep)

# Population, latest vintage
df <- get_population()

# Net migration and births, specific years
df <- get_population_change(
  variables = c("births", "net_migration"),
  years     = 2010:2024
)

# Specific counties
df <- get_population_change(
  variables = c("domestic_migration", "international_migration"),
  geoids    = c("54011", "54025")
)

# Variable documentation
get_pep_codebook()
```

## Key variables

| Variable | Coverage | Description |
|---|---|---|
| `population` | 2000–present | Total resident population |
| `pop_16plus` | 2007–present | Resident population age 16+ |
| `births` | 2000–present | Annual births |
| `deaths` | 2000–present | Annual deaths |
| `natural_change` | 2000–present | Natural change (births minus deaths) |
| `domestic_migration` | 2000–present | Net domestic migration |
| `international_migration` | 2000–present | Net international migration |
| `net_migration` | 2000–present | Net migration (domestic + international) |

Use `get_pep_codebook()` for the current, authoritative variable list —
package documentation has shown short forms (`net_mig`, `domestic_mig`)
in one place and these full names in working code examples elsewhere;
the names above are what the package's own vignettes actually call.

### Multi-vintage coverage

Three Census PEP vintages are stitched together internally:

| Vintage | Source |
|---|---|
| 2000–2009 | Intercensal county estimates |
| 2010–2019 | Vintage 2019 county estimates |
| 2020–present | Latest vintage county estimates |

## Core functions

`get_population()` returns total population; `get_population_change()`
returns the components-of-change variables:

```{r}
#| eval: false

# Total population, all counties, all years
pop <- get_population(geography = "county", variables = "population")

# Migration components for West Virginia counties
wv_mig <- get_population_change(
  variables = c("domestic_migration", "international_migration", "net_migration"),
  geoids    = grep("^54", unique(pop$geoid), value = TRUE),
  years     = 2010:2025
)
```

## Common use cases

### Components of population change

Population change is driven by two forces: natural change (births
minus deaths) and migration (domestic and international):

```{r}
#| eval: false

library(dplyr)

grafton <- get_population_change(
  variables = c("natural_change", "domestic_migration", "international_migration"),
  geoids    = "33009",
  years     = 2000:2025
)
```

CORI's own analysis of Grafton County, NH found migration patterns —
not natural change — are what shape the county's population
trajectory.

### Rural vs. nonrural population growth

Indexing population to a base year isolates relative growth trends
between rural and nonrural counties:

```{r}
#| eval: false

library(ruraldefinitions)

rural_definition <- cbsa_2023 |> select(geoid, is_rural)

pop <- get_population(variables = "population", years = 2000:2025) |>
  left_join(rural_definition, by = "geoid") |>
  filter(!is.na(is_rural)) |>
  group_by(year, is_rural) |>
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") |>
  group_by(is_rural) |>
  mutate(index = value / value[year == 2000] * 100) |>
  ungroup()
```

CORI's own analysis using this pattern found rural population growth
leveled off after the Great Recession, while nonrural growth continued.

### Natural change vs. migration, rural counties

Aggregating across all rural counties separates the two drivers of
population change — useful because they imply different policy levers:

```{r}
#| eval: false

rural_comp <- get_population_change(
  variables = c("natural_change", "net_migration"),
  years     = 2000:2025
) |>
  left_join(rural_definition, by = "geoid") |>
  filter(is_rural == "Rural") |>
  group_by(year, variable) |>
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop")
```

## Data coverage

| Dimension | Coverage |
|---|---|
| **Geography** | County (5-digit FIPS) |
| **Time period** | 2000–present |
| **Update frequency** | Annual (Census releases new vintage each year) |
| **Source** | Census Bureau Population Estimates Program |

## Use cases

-   **Population trend analysis** — Track rural population growth and
    decline from 2000 to present
-   **Migration decomposition** — Separate domestic and international
    migration to understand drivers of population change
-   **Natural change analysis** — Identify counties where deaths exceed
    births (natural population decline)
-   **Per-capita denominators** — Provide population base for rate
    calculations in other CORI analyses
-   **Workforce proxies** — Use `pop_16plus` as a working-age population
    denominator for employment rate calculations

## Related resources

### Datasets

-   [Census Population Estimates](/datasets/census-population-estimates/) —
    Source dataset documentation (population totals)
-   [Census PEP Components of Change](/datasets/census-pep-components/) —
    Source dataset documentation (births, deaths, migration)

### Related packages

-   [ruraldefinitions](/packages/ruraldefinitions/) — Rural/nonrural
    county classification for stratified population analysis

### External links

-   [Census PEP Homepage](https://www.census.gov/programs-surveys/popest.html)
-   [Vintage 2024 County Data](https://www.census.gov/data/tables/time-series/demo/popest/2020s-counties-total.html)
-   [PEP Methodology](https://www.census.gov/programs-surveys/popest/technical-documentation/methodology.html)

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data.pep)
-   [Full README](https://github.com/ruralinnovation/cori.data.pep/blob/main/README.md)
-   [Vignette: Introduction](https://github.com/ruralinnovation/cori.data.pep/blob/main/vignettes/introduction.Rmd)
-   [Vignette: Rural Population Trends](https://github.com/ruralinnovation/cori.data.pep/blob/main/vignettes/rural-population-trends.Rmd)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.pep/issues).
