---
title: "cori.data.pep"
description: "Access and analyze U.S. Census Bureau Population Estimates Program (PEP) data at the county level from 2000 to present, including population totals and components of change"
author: "Drew Rosebush"
date: "2026-05-12"
categories: ["R Package", "Demographics"]
tags: ["PEP", "Population", "Migration", "County-Level", "S3", "Tidy Data", "Rural"]
package:
  name: "cori.data.pep"
  githubUrl: "https://github.com/ruralinnovation/cori.data.pep"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.pep")'
  status: "development"
  version: "0.1.0"
  maintainer: "Drew Rosebush"
  featured: false
usesDatasets: ["census-population-estimates", "census-pep-components"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.pep` package provides access to U.S. Census Bureau
Population Estimates Program (PEP) data at the county level from 2000
to present. It covers eight variables including total population,
working-age population (16+), and a full decomposition of population
change into natural change (births minus deaths) and net migration
(domestic and international). Data span three Census vintages, handled
internally by the package.

::: callout-note
## Why Use This Package

-   **Eight variables** — Total population, pop 16+, births, deaths,
    natural change, domestic migration, international migration, net
    migration
-   **Continuous panel from 2000** — Spans three decennial vintages
    (2000–2009, 2010–2019, 2020–present), stitched into a single
    consistent series
-   **S3-backed access** — Pre-processed data available via DuckDB for
    fast county-level queries
-   **Standard long format** — One row per `geoid / year / variable`,
    ready for tidy analysis
:::

## Installation

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.pep")
```

## Quick Start

```{r}
#| eval: false

library(cori.data.pep)

# All data, latest vintage
df <- read_pep_from_s3()

# Total population and migration, specific years
df <- read_pep_from_s3(
  variables = c("population", "net_mig"),
  years     = 2010:2024
)

# Specific counties
df <- read_pep_from_s3(
  variables = c("domestic_mig", "international_mig"),
  geoids    = c("54011", "54025")
)

# Variable documentation
get_pep_codebook()
```

## Key Features

### Eight Population and Migration Variables

| Variable | Coverage | Description |
|---|---|---|
| `population` | 2000–present | Total resident population |
| `pop_16plus` | 2007–present | Resident population age 16+ |
| `births` | 2000–present | Annual births |
| `deaths` | 2000–present | Annual deaths |
| `natural_chg` | 2000–present | Natural change (births minus deaths) |
| `domestic_mig` | 2000–present | Net domestic migration |
| `international_mig` | 2000–present | Net international migration |
| `net_mig` | 2000–present | Net migration (domestic + international) |

### Multi-Vintage Coverage

Three Census PEP vintages are handled internally:

| Vintage | URL Source |
|---|---|
| 2000–2009 | Intercensal county estimates |
| 2010–2019 | Vintage 2019 county estimates |
| 2020–present | Latest vintage county estimates |

### Data Refresh (Admin)

After Census releases new estimates:

```{r}
#| eval: false

write_pep_processed_to_s3(years = 2000:2025)
```

## Core Functions

| Function | Category | Description |
|---|---|---|
| `read_pep_from_s3()` | Data Access | Read processed PEP data from S3 via DuckDB |
| `pull_population()` | Data Access | Pull population totals from Census source files |
| `pull_pop_16plus()` | Data Access | Pull population age 16+ from Census source files |
| `pull_components()` | Data Access | Pull components of change from Census source files |
| `write_pep_processed_to_s3()` | Admin | Process and write PEP data to S3 |
| `get_pep_codebook()` | Utility | Return variable definitions and metadata |
| `latest_pep_vintage()` | Utility | Return the most recent available vintage year |

## Data Coverage

| Dimension | Coverage |
|---|---|
| **Geography** | County (5-digit FIPS) |
| **Time Period** | 2000–present |
| **Update Frequency** | Annual (Census releases new vintage each year) |
| **Source** | Census Bureau Population Estimates Program |

## Use Cases

-   **Population trend analysis** — Track rural population growth and
    decline from 2000 to present
-   **Migration decomposition** — Separate domestic and international
    migration to understand drivers of population change
-   **Natural change analysis** — Identify counties where deaths exceed
    births (natural population decline)
-   **Per-capita denominators** — Provide population base for rate
    calculations in other CORI analyses
-   **Workforce proxies** — Use `pop_16plus` as a working-age population
    denominator for employment rate calculations

## Related Resources

### Datasets

-   [Census Population Estimates](/datasets/census-population-estimates/) —
    Source dataset documentation (population totals)
-   [Census PEP Components of Change](/datasets/census-pep-components/) —
    Source dataset documentation (births, deaths, migration)

### Related Packages

-   [ruraldefinitions](/packages/ruraldefinitions/) — Rural/nonrural
    county classification for stratified population analysis

### External Links

-   [Census PEP Homepage](https://www.census.gov/programs-surveys/popest.html)
-   [Vintage 2024 County Data](https://www.census.gov/data/tables/time-series/demo/popest/2020s-counties-total.html)
-   [PEP Methodology](https://www.census.gov/programs-surveys/popest/technical-documentation/methodology.html)

## Documentation

-   [GitHub Repository](https://github.com/ruralinnovation/cori.data.pep)
-   [Full README](https://github.com/ruralinnovation/cori.data.pep/blob/main/README.md)
-   [Vignette: Introduction](https://github.com/ruralinnovation/cori.data.pep/blob/main/vignettes/introduction.Rmd)
-   [Vignette: Rural Population Trends](https://github.com/ruralinnovation/cori.data.pep/blob/main/vignettes/rural-population-trends.Rmd)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.pep/issues).
