---
title: "cori.data.bds"
description: "Access and analyze Census Bureau Business Dynamics Statistics (BDS) data covering establishment and job dynamics from 1978-2023"
author: "Drew Rosebush"
date: "2026-08-26"
categories: ["R Package", "Entrepreneurship"]
tags: ["Census", "Business Dynamics", "Employment", "Startups", "Tidy Data", "Economic Development"]
package:
  name: "cori.data.bds"
  githubUrl: "https://github.com/ruralinnovation/cori.data.bds"
  installCommand: 'remotes::install_github("ruralinnovation/cori.data.bds")'
  status: "stable"
  version: "0.3.0"
  maintainer: "Drew Rosebush"
  featured: true
usesDatasets: ["census-bds"]
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.bds` package provides tools for accessing and analyzing Census Bureau Business Dynamics Statistics (BDS) data. This dataset tracks establishment and job dynamics from 1978 to the present — not just how many jobs exist, but where those jobs came from: new businesses opening, existing businesses expanding, or businesses closing and contracting.

All `get_*` functions return **tidy (long) format** data — one row per `geoid × year × variable` combination, with metric values in a `value` column. Dimensional functions (`_by_sector`, `_by_age`, `_by_size`) add the relevant dimension column(s) as additional identifiers.

::: callout-note
## What's in the package

- **Tidy output** — every function returns long-format data, ready for `dplyr`/`tidyr` pipelines
- **Two data domains** — business dynamics (establishments, entries, exits) and employment dynamics (job creation, destruction, net change)
- **Three dimensional breakdowns** — sector, establishment age, and firm size, each mutually exclusive
- **Flexible filtering** — by geography, year, specific FIPS codes, or specific variables
:::

## Installation

::: panel-tabset
## GitHub

```{r}
#| eval: false

remotes::install_github("ruralinnovation/cori.data.bds")
```

## Requirements

**Census API key** (required):
```{r}
#| eval: false

# Get one at: http://api.census.gov/data/key_signup.html
Sys.setenv(CENSUS_API_KEY = "your_api_key_here")
```

**S3 access** (for reading processed data from S3):
```{r}
#| eval: false

# cori.data.bds depends on cori.data.s3 for S3 reads.
# Falls back to CORI's credential-vending endpoint if no
# local AWS credentials are configured.
remotes::install_github("ruralinnovation/cori.data.s3")
```
:::

## Quick start

```{r}
#| eval: false

library(cori.data.bds)

# See all available variables and which functions return them
get_bds_codebook()

# Aggregate business dynamics (no dimensional breakdown)
# Returns: geoid | year | variable | value
bd <- get_business_dynamics(geography = "county", years = 2018:2023)

# Aggregate employment dynamics
ed <- get_employment_dynamics(geography = "county", years = 2018:2023)
```

## Core functions

### Business dynamics

Track establishment births, deaths, and counts:

```{r}
#| eval: false

# Aggregate totals
get_business_dynamics(geography = "county", years = 2018:2023)

# By NAICS sector
get_business_dynamics_by_sector(geography = "state", years = 2018:2023)

# By establishment age
get_business_dynamics_by_age(geography = "county", years = 2018:2023)

# By firm size
get_business_dynamics_by_size(geography = "county", years = 2018:2023)
```

### Employment dynamics

Track job creation, destruction, and net change — dimensional breakdowns follow the same `_by_sector` / `_by_age` / `_by_size` pattern:

```{r}
#| eval: false

get_employment_dynamics(geography = "county", years = 2018:2023)
get_employment_dynamics_by_sector(geography = "state", years = 2018:2023)
get_employment_dynamics_by_age(geography = "county", years = 2018:2023)
get_employment_dynamics_by_size(geography = "county", years = 2018:2023)
```

## Dimensional breakdowns

BDS measures three **mutually exclusive** dimensions. Each `_by_*` function slices the data along one dimension at a time.

::: callout-important
## NAICS granularity differs by geography

At the **county** level, sector data is limited to the 20 major 2-digit NAICS sectors (e.g. `"23"` Construction, `"62"` Health Care). At the **state and national** level, the Census BDS API provides considerably more detail — 2, 3, and 4-digit codes across roughly 390 categories (e.g. `"2361"` Residential Building Construction, `"6221"` General Medical and Surgical Hospitals). Keep this in mind when comparing sector breakdowns across geographies.
:::

| Dimension | Groups |
|---|---|
| Sector (NAICS) | 20 major sectors (county) or ~390 codes at 2–4 digits (state/national) |
| Establishment age | 0, 1, 2, 3, 4, 5, 6–10, 11–15, 16–20, 21–25, 26+ years |
| Firm size | 1–19, 20–499, 500+ employees |

## Filtering and selection

### Filter by geography

Use `geoids` to pull specific counties or states by FIPS code:

```{r}
#| eval: false

# Specific counties
rural_counties <- get_employment_dynamics(
  geoids    = c("33009", "33011", "33013"),
  geography = "county",
  years     = 2010:2023
)

# Specific states
mountain_west <- get_business_dynamics(
  geoids    = c("30", "16", "56", "49", "08"),
  geography = "state",
  years     = 2010:2023
)
```

### Select specific variables

Use `variables` to return only the columns you need:

```{r}
#| eval: false

# Net job creation only
net_jobs <- get_employment_dynamics(
  geography = "county",
  years     = 2018:2023,
  variables = c("net_job_creation", "net_job_creation_rate")
)

# Establishment entry and exit rates only
churn <- get_business_dynamics(
  geography = "state",
  years     = 2018:2023,
  variables = c("estab_entry_rate", "estab_exit_rate")
)
```

## Key variables

### Business dynamics

| Variable | Description |
|---|---|
| establishments | Number of establishments |
| estab_entry_rate | Rate of establishment births |
| estab_exit_rate | Rate of establishment exits |
| firms | Number of firms |
| firm_deaths | Firm deaths in 12 months |

### Employment dynamics

| Variable | Description |
|---|---|
| employees | Number of employees |
| job_creation | Jobs created from births and expansions |
| job_creation_rate | Rate of job creation |
| job_destruction | Jobs lost from closures and contractions |
| job_destruction_rate | Rate of job destruction |
| net_job_creation | Net employment change |
| net_job_creation_rate | Rate of net job change |

Use `get_bds_codebook()` for complete variable documentation.

## Alternative data access

For wide-format data with original Census variable names:

```{r}
#| eval: false

# Pull specific years
raw_data <- pull_bds_raw(years = 2020:2023, geography = "state")

# Pull all available years (1978-2023)
raw_data <- pull_bds_raw(years = "all", geography = "county")

# Custom variables
raw_data <- pull_bds_raw(
  years     = 2023,
  geography = "national",
  variables = c("YEAR", "NAICS", "ESTAB", "JOB_CREATION")
)
```

## Common use cases

### Rural vs. non-rural comparison

```{r}
#| eval: false

library(dplyr)
library(ruraldefinitions)

county_ed <- get_employment_dynamics(geography = "county", years = 2018:2023)

rural_definitions <- cbsa_2023 |>
  select(geoid, is_rural)

rural_comparison <- county_ed |>
  tidyr::pivot_wider(
    id_cols     = c(geoid, year),
    values_from = "value",
    names_from  = "variable"
  ) |>
  left_join(rural_definitions, by = "geoid") |>
  group_by(year, is_rural) |>
  summarize(
    avg_job_creation_rate    = mean(job_creation_rate, na.rm = TRUE),
    avg_job_destruction_rate = mean(job_destruction_rate, na.rm = TRUE),
    avg_net_job_creation     = mean(net_job_creation_rate, na.rm = TRUE),
    .groups = "drop"
  )
```

## Data coverage

- **Time period**: 1978–2023
- **Geographies**: National, state, county
- **Sectors**: 20 major NAICS sectors (county) or ~390 codes at 2–4 digits (state/national)
- **Establishment ages**: 0, 1, 2, 3, 4, 5, 6–10, 11–15, 16–20, 21–25, 26+ years
- **Firm sizes**: 1–19, 20–499, 500+ employees
- **Update frequency**: Annual (Census releases new data each year)

## Related resources

### Datasets

- [Census Business Dynamics Statistics](/datasets/census-bds/) — Source dataset documentation
- [Census Business Formation Statistics](/datasets/census-bfs/) — Leading indicator
- [QCEW Employment & Wages](/datasets/qcew-employment-wages/) — Quarterly employment data

### External links

- [Census BDS Homepage](https://www.census.gov/programs-surveys/bds.html)
- [BDS Documentation](https://www.census.gov/programs-surveys/bds/documentation.html)
- [Census API Documentation](https://www.census.gov/data/developers/data-sets/business-dynamics.html)

## Documentation

- [GitHub Repository](https://github.com/ruralinnovation/cori.data.bds)
- [Full README](https://github.com/ruralinnovation/cori.data.bds/blob/main/README.md)
- [Introduction Vignette](https://github.com/ruralinnovation/cori.data.bds/blob/main/vignettes/introduction.Rmd)

---

**Contact:** Drew Rosebush (drew.rosebush@ruralinnovation.us)

For bug reports or feature requests, please open an issue on [GitHub](https://github.com/ruralinnovation/cori.data.bds/issues).
