---
title: cori.data.bds
description: Access and analyze Census Bureau Business Dynamics Statistics (BDS)
  data covering establishment and job dynamics from 1978 to present. Provides tidy
  access to business dynamics — entries, exits, job creation, job destruction — by
  geography, industry, establishment age, and firm size.
author: Drew Rosebush
date: '2026-08-31'
categories:
  - R Package
  - Entrepreneurship
tags:
  - BDS
  - Business Dynamics
  - Establishments
  - County-Level
  - S3
  - Tidy Data
  - Rural
package:
  name: cori.data.bds
  githubUrl: https://github.com/ruralinnovation/cori.data.bds
  installCommand: remotes::install_github("ruralinnovation/cori.data.bds")
  status: development
  version: 0.3.0
  maintainer: Drew Rosebush
  featured: true
usesDatasets:
  - census-bds
usesResources: []
execute:
  echo: true
  warning: false
  message: false
editor:
  markdown:
    wrap: 72
---

## Overview

The `cori.data.bds` package provides access to Census Bureau Business
Dynamics Statistics (BDS) data at the county, state, and national level
from 1978 to the present. BDS tracks establishment and job dynamics —
entries, exits, job creation, and job destruction — across dimensions
that most employment sources cannot provide: industry, establishment age,
and firm size. This makes it possible to ask not just *how many jobs
exist* in a place, but *where those jobs came from* and how the
underlying economy is churning.

All `get_*` functions return long-format data — one row per
`geoid × year × variable` — with additional dimension columns for
sector, age, and size breakdowns, consistent with the rest of the
`cori.data` ecosystem.

## Installation & Quick Start

```{r}
#| eval: false
remotes::install_github("ruralinnovation/cori.data.bds")
```

```{r}
#| eval: false
library(cori.data.bds)

# Variable definitions
get_bds_codebook()

# Aggregate business dynamics — county level
get_business_dynamics(geography = "county", years = 2018:2023)

# Aggregate employment dynamics — county level
get_employment_dynamics(geography = "county", years = 2018:2023)

# By NAICS sector — state level
get_business_dynamics_by_sector(geography = "state", years = 2018:2023)

# By establishment age — county level
get_business_dynamics_by_age(geography = "county", years = 2018:2023)
```

## What This Data Tells Us About Rural

Business dynamics data captures what employment data misses: the
churning process beneath aggregate counts. A county can hold steady
employment while simultaneously losing established firms and gaining new
ones — or vice versa. Entry and exit rates, job creation by firm age,
and establishment counts together describe whether a rural economy is
renewing itself or sustaining on its existing institutional base.

Young firms — those five years old or less — create a disproportionate
share of net new jobs in the U.S. economy. Rural communities with active
entrepreneurial ecosystems show this pattern clearly, and CORI's
community research documents meaningful business formation growth in
communities that have invested in their entrepreneurial infrastructure.
Converting new business formation into employment growth takes time,
capital, and ecosystem depth — the BDS data supports tracking that
progression over multiple years rather than treating any single year as
definitive.

High exit rates in isolation can look like distress. But exit rates
interpreted alongside entry rates tell a more nuanced story: dynamic
economies churn. A community with both high entry and high exit rates may
be more economically active than one with low rates of both. The BDS
data supports this kind of comparative reading across rural county types
and over time.

Industry concentration at the county level reveals the degree to which
a rural economy's employment base is anchored in a narrow set of sectors
— often a direct reflection of the community's core assets.
Understanding that concentration is the starting point for any economic
development strategy, whether the goal is to deepen existing strengths or
to build complementary capacity alongside them.

## Introduction

*Adapted from the [Introduction
vignette](https://github.com/ruralinnovation/cori.data.bds/blob/main/vignettes/introduction.Rmd)
on GitHub.*

### What is this data?

The Census Bureau's Business Dynamics Statistics track the birth, death,
and employment of U.S. establishments and firms from 1978 to the
present. Unlike most employment data sources, BDS lets you ask not just
*how many jobs* exist, but *where those jobs came from* — new businesses
opening, existing businesses expanding, or businesses closing and
contracting.

Data is available at the county, state, and national level.

### Quick start

```{r}
#| eval: false
library(cori.data.bds)

get_bds_codebook()

# Aggregate — no dimensional breakdown
bd <- get_business_dynamics(geography = "county", years = 2018:2023)
ed <- get_employment_dynamics(geography = "county", years = 2018:2023)
```

### Dimensional breakdowns

BDS measures three mutually exclusive dimensions. Each `_by_*` function
slices the data along one dimension at a time.

```{r}
#| eval: false
# By NAICS sector
bd_sector <- get_business_dynamics_by_sector(geography = "state", years = 2018:2023)
ed_sector <- get_employment_dynamics_by_sector(geography = "state", years = 2018:2023)
```

**Note on NAICS granularity:** At the county level, data is limited to
20 major 2-digit NAICS sectors. At state and national level, the Census
API provides considerably more detail — 2-, 3-, and 4-digit codes across
approximately 390 categories.

```{r}
#| eval: false
# By establishment age
bd_age <- get_business_dynamics_by_age(geography = "county", years = 2018:2023)
ed_age <- get_employment_dynamics_by_age(geography = "county", years = 2018:2023)
```

```{r}
#| eval: false
# By firm size (mutually exclusive groups: 1-19, 20-499, 500+)
bd_size <- get_business_dynamics_by_size(geography = "county", years = 2018:2023)
ed_size <- get_employment_dynamics_by_size(geography = "county", years = 2018:2023)
```

### Filtering by geography and variable

```{r}
#| eval: false
# Specific counties
rural_counties <- get_employment_dynamics(
  geoids = c("33009", "33011", "33013"),
  years  = 2010:2023
)

# Specific variables only
net_jobs <- get_employment_dynamics(
  geography = "county",
  years     = 2018:2023,
  variables = c("net_job_creation", "net_job_creation_rate")
)
```

### Rural vs. nonrural comparison

```{r}
#| eval: false
library(dplyr)
library(ruraldefinitions)

county_ed <- get_employment_dynamics(geography = "county", years = 2018:2023)

county_ed |>
  left_join(select(cbsa_2023, geoid, is_rural), by = "geoid") |>
  filter(!is.na(is_rural)) |>
  group_by(year, is_rural) |>
  summarize(
    avg_job_creation_rate    = mean(job_creation_rate, na.rm = TRUE),
    avg_job_destruction_rate = mean(job_destruction_rate, na.rm = TRUE),
    avg_net_job_creation     = mean(net_job_creation_rate, na.rm = TRUE),
    .groups = "drop"
  )
```

## Related

-   **Dataset:** [Census Business Dynamics
    Statistics](/datasets/census-bds/)
-   **Package:** [cori.data](/packages/cori-data/) — meta-package that
    includes `cori.data.bds`
-   **Package:** [cori.data.bfs](/packages/cori-data-bfs/) — business
    applications as a leading indicator; pairs with BDS for the full
    formation-to-employment story
-   **Package:** [ruraldefinitions](/packages/ruraldefinitions/) —
    rural/nonrural classification used in comparison examples
-   **Project:** [Rural Economic
    Outlook](/projects/rural-economic-outlook/) — BDS is a core input to
    CORI's multi-dataset economic analysis

::: callout-note
**Data attribution:** Business Dynamics Statistics are produced by the
U.S. Census Bureau. When citing analysis using this package, reference:
U.S. Census Bureau, Business Dynamics Statistics,
[www.census.gov/programs-surveys/bds](https://www.census.gov/programs-surveys/bds).
:::
